### 2017-04-06 stats3.tex
Latex conversion

\chapter*{Statistics 3}

\section{Contingency Tables}

    \newpage
    \subsection{Contingency Tables}
        So far we have looked at the frequency a single event happens, but sometimes your interested in the frequencies of two criteria happening.

        \begin{example}
        {
            Trains from 3 stations leave late or on-time, test at 5\% is their is any association.\\

            \begin{center}
            \begin{tabular}{c|c|c}
                  & On Time & Late \\
                \hline
                A & 26      & 14   \\
                B & 30      & 10   \\
                C & 44      & 26   \\
            \end{tabular}
            \end{center}
        }

        \begin{step}{Hypothesis}
        $H_0$: No association

        $H_1$: Association
        \end{step}

        \begin{step}{Calculate the expected and $\chi^2$ values}
        \begin{center}
        \begin{tabular}{c|c|c}
            $O_i$ & $E_i$           & $\displaystyle\frac{(O_i - E_i)^2}{E_i}$   \\[2ex]
            \hline
            \rule{0pt}{3.5ex} 
            26    & $\frac{80}{3}$  & $\frac{1}{60}$                \\[1ex]
            30    & $\frac{80}{3}$  & $\frac{5}{12}$                \\[1ex]
            44    & $\frac{140}{3}$ & $\frac{1}{60}$                \\[1ex]
            14    & $\frac{40}{3}$  & $0.152$                       \\[1ex]
            10    & $\frac{40}{3}$  & $\frac{1}{30}$                \\[1ex]
            16    & $\frac{70}{3}$  & $0.305$                       \\[1ex]
        \end{tabular}
        \end{center}
        \end{step}

        \begin{step}{Find the statistic $\chi^2$}
        $$\sum{\frac{(O_i - E_i)^2}{E_i}} = \chi^2 = 1.757$$
        \end{step}

        \begin{step}{Find the critical value}
        $$v = (3-1)\times(2-1) = 2$$
        $$\chi^2_2(0.95) = 5.991$$
        \end{step}

        \begin{step}{Conclusion}
        $1.757 < 5.991$ so accept $H_0$, evidence suggests stations and lateness are not associated.
        \end{step}

        \end{example}

    \newpage
    \subsection{Yates Correction}
        For contingency tables that are 2x2 we must use yate's correction since it produces a more accurate result.

        $$
        \chi^2 = \frac{(|O_i - E_i| - 0.5)^2}{E_i}
        $$

        \begin{example}
        {
            Test the results of the following drug trial at a 5\% confidence interval.

            \begin{center}
            \begin{tabular}{c|c|c}
                        & Cold  & No cold   \\
                \hline
                Drug    & 32    & 66        \\
                Placebo & 45    & 55        \\
            \end{tabular}
            \end{center}
        }

        \begin{step}{Hypothesis}
        $H_0$: No association

        $H_1$: Association
        \end{step}

        \begin{step}{Calculate the expected and $\chi^2$ values}
        \begin{center}
        \begin{tabular}{c|c|c}
            $O_i$ & $E_i$ & $\displaystyle\frac{(|O_i - E_i| - 0.5)^2}{E_i}$    \\[2ex]
            \hline
            \rule{0pt}{3.5ex} 
            34    & 36.5  & 0.633                                               \\[1ex]
            45    & 39.5  & 0.633                                               \\[1ex]
            66    & 60.5  & 0.413                                               \\[1ex]
            55    & 60.5  & 0.413                                               \\[1ex]
        \end{tabular}
        \end{center}
        \end{step}

        \begin{step}{Find the statistic $\chi^2$}
        $$\sum{\frac{(|O_i - E_i| - 0.5)^2}{E_i}} = \chi^2 = 2.09$$
        \end{step}

        \begin{step}{Find the critical value}
        $$v = (2-1)*(2-1) = 1$$
        $$\chi^2_1(0.95) = 3.841$$
        \end{step}

        \begin{step}{Conclusion}
        $2.09 < 3.841$ so accept $H_0$, no evidence to show drug is effective.
        \end{step}

        \end{example}

\newpage
\section{Distribution free methods}
    Distribution free test's are tests that do not require the data to follow a particular distribution. Their are four Distribution free methods you need to know:

    \begin{center}
    \begin{tabulary}{\linewidth}{|L|L|L|}
    \hline
    Test & Test for & When to use \\\hline
    The sign test & Median & Cant use Wilcoxon (data is not symmetrical) \\\hline
    Wilcoxon signed-rank test & Median/mean & Cant use z or t test \\\hline
    Mann-Whitney U test & Two samples have identical populations & Their are two samples \\\hline
    Kruskal-Wallis test & More than two samples from identical populations & Their are more than two samples \\\hline
    \end{tabulary}
    \end{center}

    \newpage
    \subsection{The Sign Test}

        The sign test works by checking if their is a significant difference in the median value of the samples by comparing each value pair. Unlike the Wilcoxon test the distribution does not have to be symmetrical.

        \subsubsection{Process}
            \begin{enumerate}
            \item Hypothesis.
            \item Find the sign of the difference ignoring any pairs that are equal.
            \item Count the number of positive and negative signs (these are the test statistics)
            \item Find the value of $P(X < min(a, b) | X \sim B(n, \frac{1}{2}))$ where $a$ and $b$ are the test statistics and $n$ is the number of pairs (minus equal pairs)
            \item Compare this value with the significant level, if its less than the significant level reject $H_0$
            \end{enumerate}

        \subsubsection{1-way Example}
            \begin{example}
            {
            Results from a 2005 sample of cocaine usage: 9, 26, 17, 18, 21, 16, 19, 13, 15. In 2000 the average was 14, test at the 10\% significance level if cocaine usage has risen.
            }

            \begin{step}{Hypothesis}
            $H_0$: Population median $\eta = 14$

            $H_1$: Population median $\eta > 14$
            \end{step}

            \begin{step}{Signs}
            \begin{center}
            \begin{tabular}{c|c|c|c|c|c|c|c|c|c}
                $x - 14$ & $-5$ & $12$ & $3$  & $4$  & $7$  & $2$  & $5$  & $-1$ & $1$ \\
                \hline
                Signs    & $-$  & $+$  & $+$  & $+$  & $+$  & $+$  & $+$  & $-$  & $+$  
            \end{tabular}
            \end{center}
            \end{step}

            \begin{step}{Test Statistic}
            $$2^- / 7^+$$
            \end{step}

            \begin{step}{Critical value}
            $$X \sim B(9, \frac{1}{2})$$
            $$P(X \leq 2) = 0.0898$$
            \end{step}

            \begin{step}{Conclusion}
            $0.0898 < 0.10$ so reject $H_0$, significant evidence at 10\% level to suggest median cocaine usage has increased since 2000.
            \end{step}

            \end{example}

        \subsubsection{2-way Example}
            \begin{example}
            {
            Test is their is a difference in the aerosols average effectiveness at a 5\% significance level.
            }

            \begin{step}{Hypothesis}
            $H_0$: Population median difference $\eta_d = 0$

            $H_1$: Population median difference $\eta_d \neq 0$
            \end{step}

            \begin{step}{Signs}
            \begin{center}
            \begin{tabular}{c|c|c|c|c|c|c|c|c|c}
                Patient     & 1   & 2   & 3   & 4   & 5   & 6   & 7   & 8   & 9   \\
                \hline
                Aerosol A   & 28  & 22  & 10  & 40  & 18  & 52  & 49  & 40  & 34  \\
                Aerosol B   & 24  & 16  & 5   & 17  & 23  & 57  & 30  & 16  & 14  \\
                Sign        & $+$ & $+$ & $+$ & $+$ & $-$ & $-$ & $+$ & $+$ & $+$ \\
            \end{tabular}
            \end{center}
            \end{step}

            \begin{step}{Test Statistic}
            $$2^- / 7^+$$
            \end{step}

            \begin{step}{Critical value}
            $$X \sim B(9, \frac{1}{2})$$

            $$P(X \leq 2) = 0.0898$$
            \end{step}

            \begin{step}{Conclusion}
            $0.0898 > 0.05$ so accept $H_0$, significant evidence at 5\% level to suggest their is no difference in the average effectiveness of aerosols.
            \end{step}

            \end{example}

    \newpage
    \subsection{Wilcoxon Signed-rank Test}

        Wilcoxon signed-rank test is also a distribution free test, for when we can't use a z or t test. Its similar to the sign test except we rank the differences \(ignoring the signs, smallest first\), then add the ranks with the same signs.

        \subsubsection{Process}
            \begin{enumerate}
                \item Hypothesis
                \item Rank the absolutes of the differences.
                \item Give the ranks the same sign as the difference.
                \item Sum the positive then the negative ranks and pick the smallest value.
                \item Find the critical value in the table.
                \item If the statistic is less than the critical value reject $H_0$.
            \end{enumerate}

        \begin{example}
        {
            Test if the distributions of mock and a-level results are the same.
        }

        \begin{step}{Hypothesis}
        $H_0$: Population median difference $\eta_d = 0$

        $H_1$: Population median difference $\eta_d > 0$
        \end{step}

        \begin{step}{Find the difference, rank and signed rank}
        \begin{center}
        \begin{tabular}{c|c|c|c|c|c}
            Candidate & Mock & A level & Difference & Rank & Signed rank \\
            1         & 40   & 45      & 5          & 7    & 7           \\
            2         & 65   & 68      & 3          & 4    & 4           \\
            3         & 53   & 47      & -6         & 9.5  & -9.5        \\
            4         & 79   & 75      & -4         & 6    & -6          \\
            5         & 87   & 88      & 1          & 1    & 1           \\
            6         & 70   & 88      & 18         & 13   & 13          \\
            7         & 80   & 77      & -3         & 4    & -4          \\
            8         & 63   & 69      & 6          & 9.5  & 9.5         \\
            9         & 51   & 60      & 9          & 12   & 12          \\
            10        & 82   & 88      & 6          & 9.5  & 9.5         \\
            11        & 27   & 30      & 3          & 4    & 4           \\
            12        & 71   & 73      & 2          & 2    & 2           \\
            13        & 29   & 35      & 6          & 9.5  & 9.5         \\
        \end{tabular}
        \end{center}
        \end{step}

        \begin{step}{Test statistics}
        \begin{align*}
        T_+ &= \sum{+ \text{Rank}} = 71.5\\
        T_- &= \sum{- \text{Rank}} = 19.5\\
        T &= 19.5\\
        \end{align*}
        \end{step}

        \begin{step}{Critical value}
        $$\text{Critical value} = 21$$
        \end{step}

        \begin{step}{Conclusion}
        $19.5 < 21$ so reject $H_0$, evidence students did better in their A-levels.
        \end{step}
        \end{example}

    \newpage
    \subsection{Mann-Whitney U-Test}

        Mann-Whitney U-Test is another non-parametric test that is used to test if two sample were taken from the same population. It is used when we can do a t-test because the data is not normal.

        \subsubsection{Process}
            \begin{enumerate}
                \item Hypothesis
                \item Rank the data (as if it was just a single set) and calculate the sum of the ranks of each set.
                \item Calculate the statistic for each set using $U = T - \displaystyle\frac{n(n+1)}{2}$ where $T$ is the sum of the ranks and $n$ is the sample size. The test statistic is the minimum of the two $U$ values.
                \item Find the critical value in the table.
                \item If the statistic is less than the critical value reject $H_0$.
            \end{enumerate}

        \begin{example}
        {
        Nine random plants from two sides of a valley where weighed, carry out a Mann-Whitney U test at 5\% level of significance.

        \begin{center}
        \begin{tabular}{c|c|c|c|c|c|c|c|c|c}
            East Side & $$27.1$$ & $$40.3$$ & $$15.7$$ & $$36.4$$ & $$16.3$$ & $$15.3$$ & $$32.0$$ & $$15.7$$ & $$27.5$$ \\
            \hline
            West Side & $$11.7$$ & $$14.7$$ & $$19.1$$ & $$22.0$$ & $$6.7$$ & $$14.1$$ & $$20.1$$ & $$24.4$$ & $$15.4$$
        \end{tabular}
        \end{center}
        }

        \begin{step}{Hypothesis}
        $H_0$: Samples taken from identical populations

        $H_1$: Samples not taken from identical populations
        \end{step}

        \begin{step}{Rank the data}
        \begin{center}
        \begin{tabular}{c|c|c|c|c|c|c|c|c|c}
        East Side & $14$ & $18$ & $7.5$ & $17$ & $9$ & $5$ & $16$ & $7.5$ & $15$  \\
        \hline
        West side & $2$  & $4$  & $10$  & $12$ & $1$ & $3$ & $11$ & $13$  & $6$   \\
        \end{tabular}
        \end{center}

        \begin{align*}
        T_E &= 14 + 18 + ... + 15 = 109\\
        T_W &= 2 + 5 + ... + 6 = 62
        \end{align*}
        \end{step}

        \begin{step}{Test Statistic}
        \begin{align*}
        U_E &= 109 - \frac{9 \times (9 + 1)}{2} = 64\\
        U_W &= 62 - \frac{9 \times (9 + 1)}{2} = 17\\
        U &= min(64, 17) = 17
        \end{align*}
        \end{step}

        \begin{step}{Critical value}
        $$\text{Critical value} = 18$$
        \end{step}

        \begin{step}{Conclusion}
        $17 < 18$ so reject $H_0$, the population average weight differs across each side.
        \end{step}
        \end{example}

    \newpage
    \subsection{Kruskal-Wallis Test}
        Kruskal-Wallis test is a non-parametric version of an ANOVA test. The test determins if their is a difference between some samples. We use this test instead of an ANOVA test if the data is not normally distributed.

        \subsubsection{Process}
            \begin{enumerate}
            \item Hypothesis
            \item Rank the data (as if it was a single set) and calculate the sum of the ranks of each set.
            \item Calculate the test statistic using: $H = \displaystyle\frac{12}{N(N + 1)} \times \displaystyle\sum{\frac{T_i^2}{n_i}} - 3(N + 1)$ where $T_i$ is the sum of all ranks in a set of size $n_i$ and $N$ is the sum of all sample sizes.
            \item Find the degrees of freedom, which is the number of samples minus one.
            \item Use the $\chi^2$ distribution to find the critical value.
            \item Compare the test statistic and critical value, if statistic is larger than critical value reject $H_0$.
            \end{enumerate}

        \begin{example}
        {
            Carry out a Kruskal-Wallis test to see if their is a difference between the fish prices from three different markets at 5\% significance level.

            \begin{center}
            \begin{tabular}{c|c|c}
                A     & B     & C     \\
                \hline
                220.3 & 190.1 & 228.7 \\
                226.3 & 209.7 & 231.3 \\
                227.3 & 223.4 & 249.6 \\
                228.1 & 224.2 & 260.7 \\
                242.4 & 226.4 & 289.7
            \end{tabular}
            \end{center}
        }

        \begin{step}{Hypothesis}
        $H_0$: Samples from the same population

        $H_1$: Samples from the different populations
        \end{step}

        \begin{step}{Rank the data}
        \begin{center}
        \begin{tabular}{c|c|c}
        A  & B & C  \\
        \hline
        3  & 1 & 10 \\
        6  & 2 & 11 \\
        8  & 4 & 13 \\
        9  & 5 & 14 \\
        12 & 7 & 15
        \end{tabular}
        \end{center}
        \end{step}

        \begin{step}{Test statistic}
        \begin{align*}
        T_A &= 3 + 6 + ... + 12 = 38\\
        T_B &= 1 + 2 + ... + 7 = 19\\
        T_C &= 10 + 11 + ... + 15 = 63\\
        \end{align*}

        \begin{align*}
        \sum\frac{T_i}{n_i} &= \frac{38^2}{5} + \frac{19^2}{5} + \frac{63^2}{5}\\
        &= 1154.8
        \end{align*}

        \begin{align*}
        H &= \frac{12}{15(15 + 1)} \times 1154.8 - 3(15 + 1)\\
        &= 9.74
        \end{align*}
        \end{step}

        \begin{step}{Degrees of freedom}
        \begin{align*}
        v &= 3 - 1 \\
          &= 2
        \end{align*}
        \end{step}

        \begin{step}{Critical value}
        $$
        \chi^2_2(5\%) = 5.991
        $$
        \end{step}

        \begin{step}{Conclusion}
        $9.74 > 5.991$ so reject $H_0$, samples are not from identical populations
        \end{step}
        \end{example}

\newpage
\section{Correlation}
    \newpage
    \subsection{Spearman's Rank Correlation Coefficent}

        The product moment correlation coefficient works when you have something to measure. Suppose you had an order, for example an order of preference in different blends of tea, this is when we use Spearman's rank correlation coefficient.

        To calculate the Spearman's rank correlation coefficient we use the following:
        $$
        r_s = 1 - \frac{6\sum{d^2}}{n(n^2-1)}
        $$
        Where $d_i$ is the difference between rank $x_i$ and rank $y_i$ and $n$, the number of pairs.

        \begin{example}
        {
            Two judges judge a competition, comment on how well they agree

            \begin{center}
            \begin{tabular}{l|c|c|c|c|c|c|c|c|c|c}
            Competitor & A & B & C & D & E & F & G & H & I & J \\
            \hline
            Judge 1 & $7.8$ & $6.6$ & $7.3$ & $7.4$ & $8.4$ & $6.5$ & $8.9$ & $8.5$ & $6.7$ & $7.7$ \\
            Judge 2 & $8.1$ & $6.8$ & $8.2$ & $7.5$ & $8.0$ & $6.7$ & $8.5$ & $8.3$ & $6.6$ & $7.8$ \\
            \end{tabular}
            \end{center}
        }

        \begin{step}{Rank the scores and find the difference}
        \begin{center}
        \begin{tabular}{l|c|c|c|c|c|c|c|c|c|c}
        Competitor & A & B & C & D & E & F & G & H & I & J \\
        \hline
        $x_i$ & 4 & 9 & 7 & 6 & 3 & 10 & 1 & 2 & 8 & 5 \\
        $y_i$ & 4 & 8 & 3 & 7 & 5 & 9 & 1 & 2 & 10 & 6 \\
        $d_i$ & 0 & 1 & 4 & -1 & -2 & 1 & 0 & 0 & -2 & -1 \\
        \end{tabular}
        \end{center}
        \end{step}

        \begin{step}{Calculate to coefficient}
        \begin{align*}
        r_s &= 1 - \frac{6 \times (0^2 + 2^2 +4^2 + ...)}{10 \times (10^2-1)}\\
        &= 1 - \frac{168}{990}\\
        &= 0.830
        \end{align*}
        \end{step}

        \begin{step}{Conclusion}
        Fairly strong correlation between the judges, suggesting judges have similar criteria and standards.
        \end{step}

        \end{example}

    \newpage
    \subsection{Testing the correction coefficient}
        We may be interested in conducting a test to show their is no correlation between two random variables. First we find the correlation coefficient, then get the critical value from the appropriate table.

        \begin{example}
        {
            Test to see if their is no correlation between English and maths marks at the 5\% significance level.

            \begin{center}
            \begin{tabular}{l|c|c|c|c|c|c|c|c}
            Student        & A  & B  & C  & D  & E  & F  & G  & H  \\
            English        & 25 & 18 & 32 & 27 & 21 & 35 & 28 & 30 \\
            Mathematics    & 16 & 11 & 20 & 17 & 15 & 26 & 32 & 20 \\
            \end{tabular}
            \end{center}
        }

        \begin{step}{Hypothesis} 
        $$H_0: \rho = 0$$

        $$H_1: \rho \ne 0$$
        \end{step}

        \begin{step}{Calculate the product moment correlation coefficient}
        \begin{align*}
        \sum{x_i} &= 216\\
        \sum{y_i} &= 157\\
        \sum{x_i^2} &= 6052\\
        \sum{y_i^2} &= 3391\\
        \sum{x_i y_i} &= 4418\\
        S_{xx} &= 6052 - \frac{216^2}{8} = 220\\
        S_{yy} &= 3391 - \frac{157^2}{8} = 309.875\\
        S_{xy} &= 4418 - \frac{216 \times 157}{8} = 179\\
        r &= \frac{179}{\sqrt{220 \times 309.875}} = 0.686
        \end{align*}
        \end{step}

        \begin{step}{Critical value (from table)}
        $$
        \text{critical value} = \pm 0.7067
        $$
        \end{step}

        \begin{step}{Conclusion}
        $-0.7067 < 0.686 < 0.7067$ so value is not in critical region, accept $H_0$, evidence to suggest product moment correlation coefficient is zero.
        \end{step}

        \end{example}

### 2017-04-07 src/stats4.tex
Added Continuous Probability Distributions and Distributional Approximations.

\chapter{Statistic 4}

\newpage
\section{Continuouse Probability Distributions}

    \newpage
    \subsection{Combination of random varibles}
    Using the following rules from SS2, given that $X$ and $Y$ are random variables:
    $$\operatorname{E}(aX \pm bY) = a\operatorname{E}(X) \pm b\operatorname{E}(Y)$$
    and the following for $X$ and $Y$ are independent random variables
    $$
    \operatorname{Var}(aX \pm bY) = a^2\operatorname{Var}(X) + b^2\operatorname{Var}(Y)
    $$
    It follows that, if $X$ and $Y$ are normally distributed:
    \begin{align*}
    aX \pm bY &\sim \operatorname{N}(E(aX \pm bX), Var(aX \pm bX))\\
    &\sim \operatorname{N}(a\operatorname{E}(X) \pm b\operatorname{E}(Y), a^2\operatorname{Var}(X) + b^2\operatorname{Var}(Y))\\
    &\sim N(a \mu_X \pm b \mu_Y, a^2 \sigma_X^2 + b^2 \sigma_Y^2)
    \end{align*}

\newpage
\section{Distributional Approximations}

    \newpage
    \subsection{Continuity corrections}
        When we go from a discrete distribution to a continuous distribution, we must account for the discrete nature of the problem in are calculations. For example, let $X$ be the number of heads after 10 flips of a coin, find the $P(X = 4)$. If we approximate the value with a continuose distribution we must find the area under the curve between $0.35$ and $0.45$ since these are the lower and upper limits of $4$.

    \newpage
    \subsection{Approximating binomial using the Possion distribution}
        If $n$ is large and $p$ is small the binomial distribution can be approximated by the Possion distribution. Good approximation can also be made with small values of $n$ given that $p$ is small enough.

        \paragraph{Conditions}
        \begin{itemize}
            \item $n > 50$
            \item $p < 0.1$
        \end{itemize}

        \begin{example}
        {
        Based upon past experience, 1\% of the telephone bills mailed to households are incorrect. If a sample of 20 bills is selected, find the probability that at least one bill will be incorrect.
        }

        \begin{step}{Find the mean}
        \begin{align*}
        \mu &= \lambda\\
        &= np\\
        &= 20 * 0.01\\
        &= 0.2
        \end{align*}
        \end{step}

        \begin{step}{Approximate using Poisson}
        \begin{align*}
        X &\sim Po(0.2)\\
        P(X \geq 1) &= P(X \leq 0)\\
        &= P(X = 0)\\
        &= \frac{e^{-0.2} \times \lambda^0}{0!}\\
        &= 0.819
        \end{align*}
        \end{step}

        \end{example}

    \newpage
    \subsection{Approximating binomial using the normal distribution}
        Central limit therom states the means of random samples are approximately normally distributed given $n$ is large enough. The total number of successes for a binomial distribution is the mean multiplied by the number of trails hence the normal distribution can be uses as an approximation for the binomal distribution.

        \paragraph{Conditions}
        \begin{itemize}
            \item $n > 50$
            \item $np > 10$
            \item $p \leq 0.5$ (invert the definition of success if this is not true)
        \end{itemize}

        \begin{example}
        {
        Suppose that a sample of $n = 1600$ tires of the same type are obtained at random from an ongoing production process in which 8\% of all such tires produced are defective. What is the probability that in such a sample 150 or fewer tires will be defective?
        }

        \begin{step}{Find mean and varience}
        \begin{align*}
        \mu &= np\\
        &= 1600 \times 0.08 = 128\\
        \end{align*}
        \begin{align*}
        \sigma^2 &= np(1-p)\\
        &= 1600 \times 0.08 \times (1 - 0.08)\\
        &= 117.76\\
        \end{align*}
        \end{step}

        \begin{step}{Aproximate using normal distribution}
        \begin{align*}
        Z &= \frac{X - \mu}{\sigma} \\
        &= \frac{150.5 - 128}{\sqrt{117.76}}\\
        &= 2.073\\
        P(Z < 2.073) &= 0.981
        \end{align*}
        \end{step}

        \end{example}

    \newpage
    \subsection{Approximating Possion using the normal distribution}
        Central limit therom may also be used to justify approximating Possion using the normal distribution given that $\lambda$ is large enough.

        \paragraph{Conditions}
        \begin{itemize}
            \item $\lambda \geq 10$
        \end{itemize}

        \begin{example}
        {
            Suppose that at a certain automobile plant the average number of work stoppages per day due to equipment problems during the production process is 12.0. What is the approximate probability of having 15 or fewer work stoppages due to equipment problems on any given day?
        }

        \begin{step}{Find the mean and variance}
        \begin{align*}
        \mu &= \lambda\\
        &= 12\\
        \end{align*}
        \begin{align*}
        \sigma^2 &= \lambda\\
        &= 12\\
        \end{align*}
        \end{step}

        \begin{step}{Aproximate using normal distribution}
        \begin{align*}
        Z &= \frac{X - \mu}{\sigma} \\
        &=\frac{15.5 - 12}{\sqrt{12}}\\
        &= 1.010\\
        P(Z < 1.010) &= 0.844\\
        \end{align*}
        \end{step}

        \end{example}

\newpage
\section{Estimation in a Real-world Context}

\newpage
\section{Application of Hypothesis Testing}

### 2017-04-07 src/stats4.tex
Added ss4 estimation.

@@
 \newpage
-\section{Continuouse Probability Distributions}
+\section{Continuous Probability Distributions}
 
@@
     \newpage
+    \subsection{Students t-distribution}
+        If the population variance is unknown due to central limit therom we know that $X$ follows a normal distribution given $n$ is large:
+        $$\frac{\bar{X} - \mu}{\displaystyle\frac{S}{\sqrt{n}}} \approx N(0, 1^2)$$
+        \\
+        However if $n$ is small we can use t-distribution to get a better estimate. 
+        $$\frac{\bar{X} - \mu}{\displaystyle\frac{S}{\sqrt{n}}} \approx t_{n-1}\text{-distribution}$$
+
+    \newpage
+    \subsection{Confidence intervals using the t-distribution}
+        If we have a small sample of values and we assume they came from a normal distribution, we can use the t-distribution to find the confidence interval for the sample using: $$ \bar{x} \pm t_{n-1}(\alpha) \times \frac{s}{\sqrt{n}} $$ where $\bar{x}$ is the sample mean, $n$ is the sample size, $\alpha$ is $\displaystyle\frac{1 - \text{significance level}}{2}$ and $s$ is the sample standard deviation.
+
+        Example
+        \begin{example}
+        {
+            Let $n = 26$, $\bar{x} = 122$ and $s^2 = 225$. Calculate a 95\% confidence interval.
+        }
+
+        \begin{step}{Solve}
+        \begin{align*} 
+        122 \pm t_{26-1}(0.025) \times \frac{\sqrt{225}}{\sqrt{26}} &= 122 \pm 2.060 \times 3.132 \\ 
+        &= (115.941, 128.059)\\
+        &= (116, 128) 
+        \end{align*}         
+        \end{step}
+
+        \end{example}
+
+    \newpage
+    \subsection{Confidence intervals for Poisson}
+        For an observation from a Poisson distribution we know that we can approximate it to a normal distribution, given $\lambda$ is large enough.
+Given this, it can be show that the approximate confidence interval for a single observation would be:
+        $$ \lambda \pm z \times \sqrt{\lambda} $$ 
+        where $z$ is the Z-value for the significance level. Note the confidence interval is an approximate since the is based of the normal distribution.
+
+        \begin{example}
+        {
+            Voters arrived between 6pm and 9pm. In the first hour 51 voters showed up. Assuing the arrivals can be moddled by a poisson distribution, calculate a 95\% confidence interval.
+        }
+
+        % TODO: create an example enviroment without steps
+        \begin{step}{Solve}
+        $$ 51 \pm 1.96 \times \sqrt{51} = (37.0, 65.0) $$
+        \end{step}
+
+        \end{example}
+
+    \newpage
+    \subsection{Confidence itervals for a proportion}
+        The parameter $p$ of a binomial distribution is the proportion of successes and its exact confidece interval can be obtained.
+Given an estimator for the proportion $\hat{p}$ we can caluate the confidence interval for the population proportion using:
+        $$\hat{p} \pm z \times \left(\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}\right)$$ 
+        where $z$ is the Z-value for the significance level
+
+        \begin{example}
+        {
+            Marzena buys Simsons matches.
+Of a random sample of 85 such matches, 27 broke when he struck them.
+Calculate an approximate 95\% confidence interval for the proportion of Simsons matches which will break when Marzena strikes them.
+        }
+
+        \begin{step}{Calculate $\hat{p}$}
+        \begin{align*}
+        \hat{p} &= \frac{27}{85} \\
+                   &= 0.31765
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Find the confidence interval}
+        \begin{align*}
+        0.31765 &\pm 1.96 \times \sqrt{\frac{0.31765(1-0.31765)}{85}} \\ 
+        0.31765 &\pm 0.09897\ (0.219, 0.417) 
+        \end{align*}
+        \end{step}
+
+        \end{example}
+
+\newpage
 \section{Application of Hypothesis Testing}

### 2017-04-07 src/stats4.tex
Added "Estimation in a Real-world Context".
Created new chapter SS5. Minor formating changes.

@@
-\chapter{Statistic 4}
+\chapter*{Statistic 4}
 
@@
 \section{Application of Hypothesis Testing}
+
+    \newpage
+    \subsection{t-distribtion hypothesis test}
+        In the same way we can use the t-distribution to find the confidence interval for a small normal sample with unknown varience, we can use the t-distribution to test the mean.
+
+        \begin{example}
+        {
+            Let $n = 20$, $\sum{x} = 21.7$, $\sum{x^2} = 28.4$. Test the mean is larger than 1.00 at a 10\% significance level.
+        }
+
+        \begin{step}{Hypothesis}
+        \begin{align*}
+        H_0&: \mu = 1.00\\
+        H_1&: \mu > 1.00
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Test statistic}
+        \begin{align*}
+        \bar{x} &= \frac{21.7}{20} = 1.085 \\
+        s &= \sqrt{\frac{1}{20 - 1} \times (28.4 - \frac{21.7^2}{20})} = 0.5055 \\ 
+        \\
+        \text{test statistic} &= \frac{1.085 - 1}{\frac{0.5055}{\sqrt{20}}} \\
+        &= 0.752
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Critical value}
+        \begin{align*}
+        t &= t_{19}(0.90) \\ 
+        &= 1.3278
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $0.752 < 1.3278$ so accept $H_0$ and reject $H_1$, their is not evidence to support the statement that the mean is larger than one.
+        \end{step}
+
+        \end{example}
+
+    \newpage
+    \subsection{Poisson hypothesis test}
+        A hypothesis test using the possion distribution is very simular previous hypothesis tests.
+
+        \paragraph{Process}
+        \begin{itemize}
+        \item Hypothesis
+        \item Find the propability that the mean is $\geq$ the sample
+        \item Compare this with significance level, if it is less than the significance level reject $H_0$
+        \end{itemize}
+
+        \begin{example}
+        {
+            Last year their was an average of 8 complaint emails per week.
+On the first week of next year 16 complaints were received.
+Test at the 5\% level, weather the mean number of complaints has increased.
+        }
+
+        \begin{step}{Hypothesis}
+        \begin{align*}
+        H_0&: \mu = 8 \\ 
+        H_1&: \mu > 8
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Test statistic}
+        $$ X \sim Po(8) $$
+        \begin{align*} 
+        P(X \geq 16) &= 1 - P(X \leq 15) \\ 
+        &= 0.00823 
+        \end{align*}         
+        \end{step}
+
+        \begin{step}{Conclusion}
+            $0.00823 < 0.05$ so reject $H_0$, the is no evidence to support the mean number of emails has increased.
+        \end{step}
+
+        \end{example}
+
+    \newpage
+    \subsection{Test for proportion}
+        The test of proportion test the population parameter $p$ (proportion) in a given sample, to access whether the sample represents the true proportion of the entire population.
+
+        \paragraph{Process}
+        \begin{itemize}
+        \item Hypothesis
+        \item Find the probability that the distribtion $B(n, p)$ is less than the sample
+        \item Compare with the significance level, if it is less than the significance level, reject $H_0$
+        \end{itemize}
+
+        \subsubsection{Example (no approximation)}
+        \begin{example}
+        {
+            40\% of passengers rated an airliners food bad.
+The airliner claimed to improve the food and re ran the survey, after which 6 out of 20 passengers said the food was bad.
+Test at a 5\% significance level if the rating the meal bad had reduced.
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0: p = 0.4$ \\
+        $H_1: p < 0.4$
+        \end{step}
+
+        \begin{step}{Probability}
+        $$X \sim B(20, 0.4)$$
+        $$P(X \leq 6) = 0.250$$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $0.250 > 0.05$ so accept $H_0$, no evidence that the proportion of passengers rating the meal bad had reduced.
+        \end{step}
+
+        \end{example}
+
+        \subsubsection{Example (with approximation)}
+        \begin{example}
+        {
+            Past experience suggests that 40\% of households donate to charity.
+In a sample 1240 households where asked if they donate, 476 had.
+Test the hypothesis that the number of households that donate is 40\%.
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0: p = 0.4$\\ 
+        $H_1: p \ne 0.4$
+        \end{step}
+
+        \begin{step}{Test statistic}
+        $$X \sim B(1240, 0.4)$$ 
+        \begin{align*}
+        \mu &= 1240 \times 0.4 = 496\\
+        \sigma^2 &= 1240 \times 0.4(1 - 0.4)\\
+        \end{align*}
+        $$X \approx N(496, 297.6)$$
+        \begin{align*}
+        z &= \frac{476.5 - 496}{\sqrt{297.6}} \\
+          &= -1.13
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Critical value}
+        $$ \text{critical value} = \pm 1.96 $$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $-1.96 < -1.13 < 1.96$ so accept $H_0$, evidence to support claim that 40\% of households make a donation.
+        \end{step}
+
+        \end{example}

### 2017-04-08 src/stats5.tex
Added "Continuous Probability Distributions" to SS5.

\chapter*{Statistics 5}

\newpage
\section{Continuous Probability Distributions}

    \newpage
    \subsection{Rectanglular Distribution}
        The rectangular distribution is a random variable that takes any value in given range within an interval, the pdf of which is:
        $$
        f(x) = 
        \begin{cases}
        \frac{1}{b - a} & a \leq x \leq b\\
        0 & \text{otherwise}
        \end{cases}
        $$
        The mean and varience are given by:
        $$\operatorname{E}(x) = \frac{b + a}{2}$$
        $$\operatorname{Var}(x) = \frac{(b - a)^2}{12}$$
        And the cumulative distribution function can be found by integrating $f(x)$ between x and the lower bound.

        \begin{example}
        {
            Given $
            f(x) = 
            \begin{cases}
                \frac{1}{6} & -2 \leq x \leq 4\\
                0 & \text{otherwise}
            \end{cases}
            $
        }

        \begin{step}{Find the mean}
        \begin{align*}
        E(x) &= \frac{4 + -2}{2} \\
             &= 1
        \end{align*}
        \end{step}

        \begin{step}{Find the varience}
        \begin{align*}
        Var(x) &= \frac{(4 - -2)^2}{12} \\
               &= 3
        \end{align*}
        \end{step}

        \begin{step}{Find the value of $P(|X| < 1)$}
        \begin{align*}
        P(|X| < 1) &= P(-1 < X < 1) \\
                   &= 2 \times \frac{1}{6} \\
                   &= \frac{1}{3}
        \end{align*}
        \end{step}

        \end{example}

    \newpage
    \subsection{Exponential Distribution}
        The exponential distribution is the interval between the successive events of a Poisson distribution. Its density function is as follows:
        $$
        f(x)
        \begin{cases}
        \lambda e^{-\lambda x} & x > 0\\
        0 & \text{otherwise}\\
        \end{cases}
        $$
        where lambda is the parameter of the exponential distribution. The mean and varience are given by:
        $$\operatorname{E}(x) = \frac{1}{\lambda}$$
        $$\operatorname{Var}(x) = \frac{1}{\lambda}$$
        If you integrate between $0$ and $x$, you can find the cumulative distribution function:
        \begin{align*}
        \int^{x}_{0}{\lambda e^{-\lambda x}}
        &= [-e^{-\lambda x}]^x_0\\
        &= [-e^{-\lambda x} -- e^{0}]^x_0\\
        &= [-e^{-\lambda x} + 1]\\
        &= 1 - e^{-\lambda x}
        \end{align*}

        \begin{example}
        {
        The number of cars passing a point per minute is 0.8. Find the probability that the interval between two cars is longer than 2 minuets.
        }

        \begin{step}{Define the density function}
        $$
        F(x) = 1 - e^{-0.8x}
        $$
        \end{step}

        \begin{step}{Find the probabilty}
        \begin{align*}
        F(\infty) - F(2) 
        &= (1 - e^{-0.8 \times \infty}) - (1 - e^{-0.8 \times 2})\\
        &= (1 - 0) - (1 - e^{-1.6})\\
        &= 0.202
        \end{align*}
        \end{step}

        \end{example}

\newpage
\section{Estimation}

    \newpage
    \subsection{Confindence interval for varience}

\newpage
\section{Hypothesis Testing}

    \newpage
    \subsection{Test for varience}

    \newpage
    \subsection{F-distribution}

    \newpage
    \subsection{Test for varience equality}

    \newpage
    \subsection{Test for mean equailty}

    \newpage
    \subsection{Goodness of fit test}

### 2017-04-08 src/stats5.tex
Added "Estimation" to SS5.

@@
     \subsection{Confindence interval for varience}
+        For a random sample of $N(\mu, \sigma^2)$ it can be shown that $\displaystyle\frac{(n-1)s^2}{\sigma^2} \sim \chi^2_{n-1}$ where $n$ is the sample size.
+Using this we can rearrange the result to find the confidence interval for varience.
+
+        \begin{example}
+        {
+            Calculate a 95\% confidence interval for the mean and variance of the set: 266, 254, 215, 220, 253, 230, 216, 248, 234, 244.
+        }
+
+        \begin{step}{Find the mean and variance}
+        \begin{align*}
+        \bar{x} &= 238 \\
+        s^2 &= 313.111
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Rearrange to find the confidence interval}
+        $$\chi^2_9(0.975) < \frac{(n-1)s^2}{\sigma^2} < \chi^2_9(0.075)$$
+        $$\chi^2_9(0.975) < \frac{9 \times 313.111}{\sigma^2} < \chi^2_9(0.075)$$
+        $$\frac{9 \times 313.111}{\chi^2_9(0.975)} < \sigma^2 < \frac{9 \times 313.111}{\chi^2_9(0.075)}$$
+        $$148.128 < \sigma^2 < 1043.553$$
+        $$(148.128, 1043.553)$$
+        \end{step}
+
+        \end{example}
 

### 2017-04-09 src/stats5.tex
Added "Hypothesis Testing" to SS5.

@@
     \subsection{Test for varience}
+        Just as we can find the confidence interval for the varience of a normal distribution, we can use the $\displaystyle\frac{(n-1)s^2}{\sigma^2}$ as a test statistic (and the $\chi^2$ for the critical value) to test a hypothesis involving the population varience.
+
+        \begin{example}
+        {
+            Some random observations (2.1, 2.3, 3.5, 4.6, 5.0, 6.4, 7.1, 8.6, 8.7, 9.1) are thought to have a variance of 4.1. Test this at a 5\% significance level.
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0: \sigma^2 = 4.1$\\
+        $H_1: \sigma^2 \ne 4.1$
+        \end{step}
+
+        \begin{step}{Test statistic}
+        \begin{align*} 
+        \bar{x} &= 5.74 \\ 
+        s^2 &= 6.940 
+        \end{align*}
+
+        $$\frac{(10 - 1) \times 6.940}{4.1} = 15.234$$
+        \end{step}
+
+        \begin{step}{Critical value}
+        \begin{align*} 
+        \chi^2_9(0.975) &= 19.023\\ 
+        \chi^2_9(0.025) &= 2.700
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $2.700 < 15.234 < 19.023$ so accept $H_0$
+        \end{step}
+
+        \end{example}
 
@@
     \subsection{F-distribution}
+        Suppose we take $n_x$ observations from a $N(\mu_x, \sigma_x^2)$ distribution and a independent sample of size $n_y$ from a $N(\mu_y, \sigma_y^2)$ distribution.\\
+        \\
+        From the previous topics, we know $$\frac{(n_x-1)s_x^2}{\sigma_x^2} \sim \chi^2_{n_x-1}$$ and $$\frac{(n_y-1)s_y^2}{\sigma_y^2} \sim \chi^2_{n_y-1}$$
+        \\
+        So if follows that: 
+        $$
+        \dfrac{\dfrac{s_x^2}{\sigma_x^2}}{\dfrac{s_y^2}{\sigma_y^2}} 
+        \sim 
+        \dfrac{\dfrac{\chi^2_{n_x-1}}{(n_x-1)}}{\dfrac{\chi^2_{n_y-1}}{(n_y-1)}}
+        $$ 
+        and 
+        $$ 
+        \dfrac{\dfrac{s_y^2}{\sigma_y^2}}{\dfrac{s_x^2}{\sigma_x^2}} 
+        \sim 
+        \dfrac{\dfrac{\chi^2_{n_y-1}}{(n_y-1)}}{\dfrac{\chi^2_{n_x-1}}{(n_x-1)}} 
+        $$
+        \\
+        This is the F-distribution and it has two parameters $(n_x - 1)$ and $(n_y - 1)$ usually written as $F_{v1, v2}$ where $v1$ and $v2$ correspond to the two parameters.\\
+        \\
+        If the parameters are the other way round (y on top) we can convert them using the following: 
+        $$ F_{v2, v1} = \frac{1}{F_{v1, v2}} $$
+
+        \begin{example}
+        {
+            $X \sim F_{8, 10}$, find $P(\frac{1}{5.81} < X < 5.06)$
+        }
+
+        \begin{step}{Solve}
+        \begin{align*} 
+        P(\frac{1}{5.81} < X < 5.06) &= P(X < 5.06) - P(X < \frac{1}{5.81})\\ 
+        &= 0.99 - P(F_{10, 8} > 5.81)\\ 
+        &= 0.99 - (1 - P(F_{10, 8} < 5.81))\\ 
+        &= 0.99 - (1 - 0.99)\\ 
+        &= 0.98 
+        \end{align*}         
+        \end{step}
+
+        \end{example}
+        \paragraph{Note}
+        To find the values on the table, look at each F-distribution table and look for the value closest to the right hand side value, in this case $5.057$ (which rounds to $5.06$) was found on the $p = 0.99$ table.
 
@@
     \subsection{Test for varience equality}
+        We can use the F-distribution to test if the two samples varience's are the same. If we assume $\sigma^2_x = \sigma^2_y$ then 
+        $$\dfrac{\dfrac{s_y^2}{\sigma_y^2}}{\dfrac{s_x^2}{\sigma_x^2}} \sim F_{n_x-1, n_y-1} $$ 
+        becomes 
+        $$\dfrac{s_y^2}{s_x^2} \sim F_{n_x-1, n_y-1}$$ 
+        which is the test statistic for this hypothesis test.
+
+        \begin{example}
+        {
+            It is believed that wood stored inside has less variable hardness than wood stored outside, conduct a test at a 0.05 significance level stating any assumptions made.
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0: \sigma_o^2 = \sigma_i^2$\\ 
+        $H_1: \sigma_o^2 > \sigma_i^2$
+        \end{step}
+
+        \begin{step}{Test statistic}
+        \begin{align*} 
+        F_{\text{test}} &= \frac{216.25}{198.6}\\ 
+        &= 1.089 
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Critical value}
+        \begin{align*} 
+        F_{\text{crit}} &= F_{25-1, 21-1}(0.05)\\ 
+        &= F_{24, 20}(0.05)\\ 
+        &= 2.08 
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $1.089 < 2.08$ so accept $H_0$, wood stored inside is just as variable as wood stored outside.
+        \end{step}
+
+        \end{example}
 
@@
 
+        \subsubsection{Two-sample z-test}
+            We may be interested in the difference between two independent populations.
+To test two distributions we can use the rules for combining distributions to derive this test statistic for two normal distributions $X$ and $Y$:
+            $$Z = \frac{\bar{X} - \bar{Y} - (\mu_x - \mu_y)} {\sqrt{\dfrac{\sigma_x^2}{n_x}+\dfrac{\sigma_y^2}{n_y}}} $$
+            If $n_x$ and $n_y$ are larger enough, this statistic can be used for distributions where $X$ and $Y$ are not normal due to central limit therom.
+
+            \begin{example}
+            {
+                A study weighs children in area $A$ and area $B$. The results are as follows:
+
+                \begin{center}
+                \begin{tabular}{l|c|c|c}
+                    & $n$ & $\bar{x}$ & $s$ \\
+                \hline
+                $A$ & 220 & 37.8      & 3.6 \\
+                $B$ & 180 & 38.6      & 4.1
+                \end{tabular}
+                \end{center}
+
+                Test (at a 5\% significance level) if their is any difference in the mean weight of children from each area.
+            }
+
+            \begin{step}{Hypothesis}
+            $H_0: \mu_A = \mu_B$ \\ 
+            $H_1: \mu_A \neq \mu_B$
+            \end{step}
+
+            \begin{step}{Test statistic} 
+            \begin{align*} 
+            z &= \frac{\bar{x}_A-\bar{x}_B - (\mu_A - \mu_B)}{\sqrt{\dfrac{\sigma_A^2}{n_A}+\dfrac{\sigma_B^2}{n_B}}}\\ 
+            &= \frac{37.6 - 38.6}{\sqrt{\dfrac{3.6^2}{220}+\dfrac{4.1^2}{180}}}\\ 
+            &= -2.56243...\\ 
+            &= -2.56 
+            \end{align*}
+            \end{step}
+
+            \begin{step}{Critical value}
+            $$z = \pm1.96$$
+            \end{step}
+
+            \begin{step}{Conclusion}
+            Since $-2.56 < -1.96$, the result is significant so reject $H_0$, their is evidence that the mean weight is different in the two areas.
+            \end{step}
+
+            \end{example}
+
+        \newpage
+        \subsubsection{Two-sample t-test}
+            In the case were the varience is unknow but equal we can use a simular test for two samples from small normal populations.
+
+            $$
+            \frac{\bar{X} - \bar{Y} - (\mu_x - \mu_y)}{\sqrt{\sigma^2\left(\dfrac{1}{n_x} + \dfrac{1}{n_y}\right)}}
+            \sim t_{n_x+n_y-2}
+            $$
+            In the case were $\sigma^2$ is unknow but the unbiased estimators ($s_x^2$ and $s_y^2$) are know, we can pool the estimators using the following formula:
+            $$
+            s_p^2 = \frac{(n_x-1)\times s_x^2 + (n_y-1) \times s_y^2}{n_x + n_y - 2}
+            $$
+
+            \begin{example}
+            {
+                Conduct a two sample t-test at a 10\% significance level with the following data:
+
+                \begin{center}
+                \begin{tabular}{l c c c c c c c c c}
+                $X$ & $40$ & $37$ & $45$ & $34$ & $30$ & $41$ & $42$ & $43$ & $36$ \\
+                \hline
+                $Y$ & $38$ & $43$ & $36$ & $45$ & $35$ & $44$ & $41$ \\
+                \end{tabular}
+                \end{center}
+            }
+
+            \begin{step}{Hypothesis}
+            $H_0: \mu_x = \mu_y$ \\ 
+            $H_1: \mu_x \ne \mu_y$ 
+            \end{step}
+
+            \begin{step}{Find the mean and varience of both samples}
+            \begin{align*}
+            n_x &= 9\\ 
+            \bar{x} &= 38.667\\ 
+            s_x^2 &= 23
+            \\
+            \\
+            n_y &= 7\\ 
+            \bar{y} &= 40.286\\ 
+            s_y^2 &= 15.905
+            \end{align*}
+            \end{step}
+
+            \begin{step}{Test statistic}
+            \begin{align*}
+            s_p^2 &= \frac{(9 - 1) \times 23 + (7 - 1) \times 15.905}{9 + 7 - 2} \\
+                &= 19.959
+            \end{align*}
+            \begin{align*}
+            t &= \frac{38.667-40.286}{\sqrt{19.959 \times \left(\dfrac{1}{9} + \dfrac{1}{7}\right)}} \\
+              &= -0.719
+            \end{align*}
+            \end{step}
+
+            \begin{step}{Critical value}
+            \begin{align*}
+            t_{9 + 7 - 2}(0.95) &= t_{14}(0.95)\\ 
+                                &= 1.761 
+            \end{align*}
+            \end{step}
+
+            \begin{step}{Conclusion}
+            $-1.761 < -0.719 < 1.761$ so accept $H_0$, evidence suggests their is no difference in the means of the two groups.
+            \end{step}
+
+            \end{example}
+
     \newpage
     \subsection{Goodness of fit test}
+        In SS3 $\displaystyle\sum{\frac{(O - E)^2}{E}}$ was used to analyse contingency tables, but we can also use this statistic to test the fit of a distribution for a particular sample.
+
+        Degrees of freedom is a measure of the data not used up.
+Every time a statistic is calculated or constraint put in place, a degree of freedom is used.
+For the purpose of testing the fit of a distribution the total frequency of the model must equal the observed frequency.
+This is an example of a constraint so we lose a degree of freedom.
+
+        \subsubsection{Discrete distributions}
+            Conditions
+            \begin{itemize}
+                \item The random varible is defined over a set of values
+                \item Each value is equally likely
+            \end{itemize}
+            The degrees of freedom for a goodness of fit test for a discrete distribution is
+            $$v = \text{number of cells} - \text{number of parameters estimated} - 1$$
+
+            \begin{example}
+            {
+                The following table shows some observed values:
+                \begin{center}
+                \begin{tabular}{l|c|c|c|c|c|c|c|c}
+                $x$ & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 \\
+                \hline
+                $O_i$ & 12 & 24 & 18 & 20 & 25 & 17 & 21 & 23 \\
+                \end{tabular}
+                \end{center}
+                Conduct a goodness of fit test for weather its a discrete uniform distribution.
+            }
+
+            \begin{step}{Hypothesis}
+            $H_0$: The observations \textit{can} be modeled by a discrete uniform distribution
+
+            $H_1$: The observations \textit{cannot} be modeled by a discrete uniform distribution
+            \end{step}
+
+            \begin{step}{Calculate the expected frequency}
+            \begin{align*}
+            E_i &= \frac{\sum{O_i}}{8} \\
+            &= \frac{160}{8} \\
+            &= 20
+            \end{align*}
+            \end{step}
+
+            \begin{step}{Test statisitic}
+            \begin{center}
+            \begin{tabular}{l|c|c|c|c|c|c|c|c}
+            $x$ & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 \\
+            \hline
+            $O_i$                                    & 12   & 24    & 18    & 20  & 25    & 17    & 21    & 23 \\
+            $E_i$                                    & 20   & 20    & 20    & 20  & 20    & 20    & 20    & 20 \\
+            $\displaystyle\frac{(O_i - E_i)^2}{E_i}$ & 3.2  & 0.8   & 0.2   & 0   & 1.25  & 0.45  & 0.05  & 0.45 \\
+            \end{tabular}
+            \end{center}
+            $$
+            \sum{\frac{(O_i - E_i)^2}{E_i}} = 6.4
+            $$
+            \end{step}
+
+            \begin{step}{Critical value}
+            \begin{align*}
+            v &= 8 - 1 \\
+            &= 7
+            \end{align*}
+            $$
+            \chi^2_7(5\%) = 14.067...
+            $$
+            \end{step}
+
+            \begin{step}{Conclusion}
+            $6.4 < 14.067...$ so accept $H_0$ and reject $H_1$, their is evidence that the observations fits a discrete uniform distribution
+            \end{step}
+            \end{example}
+
+            You can estimate the parameter $p$ of a binomial distribution with: 
+            $$ p = \frac{\text{total number of successes}}{\text{number of trials} \times \text{number of observations}} $$
+
+            \begin{example}
+            {
+                The following table shows observed and expected values for a binomial model.
+
+                \begin{center}
+                \begin{tabular}{l|c|c|c|c|c|c}
+                $O_i$ & 17 & 28 & 32 & 15 & 5 & 3 \\
+                \hline
+                $E_i$ & 19.69 & 34.74 & 27.58 & 12.98 & 4.01 & 0.99 \\
+                \end{tabular}
+                \end{center}
+
+                Test at a 5\% significance level if the distribution fits a binomial model.
+            }
+
+            \begin{step}{Hypothesis}
+            $H_0$: The observations \textit{can} be modeled by a binomial distribution
+
+            $H_1$: The observations \textit{cannot} be modeled by a binomial distribution
+            \end{step}
+
+            \begin{step}{Test statisitic}
+                Since $E_i < 5$ in the last to columns, we must combine them to work out $\chi^2$:
+
+                \begin{center}
+                \begin{tabular}{l|c|c|c|c|c}
+                $O_i$ & 17 & 28 & 32 & 15 & 9 \\
+                \hline
+                $E_i$ & 19.69 & 34.74 & 27.58 & 12.98 & 5 \\
+                $\displaystyle\frac{(O_i - E_i)^2}{E_i}$ & 0.368 & 1.31 & 0.708 & 0.314 & 1.8 \\
+                \end{tabular}
+                \end{center}
+
+                $$
+                \sum{\frac{(O_i - E_i)^2}{E_i}} = 4.50
+                $$
+            \end{step}
+
+            \begin{step}{Critical value}
+                \begin{align*}
+                v &= 5 - 1 \\
+                   &= 4
+                \end{align*}
+
+                $$
+                \chi^2_4(5\%) = 9.49...
+                $$
+            \end{step}
+
+            \begin{step}{Conclusion}
+                Since $4.50 < 9.49$ so accept $H_0$ and reject $H_1$, their is evidence to suggest the binomial model is a good fit.
+            \end{step}
+
+        \end{example}
+
+        You can estimate $\lambda$ with: $$ \lambda = \frac{\sum{r \times f_r}}{N} $$
+
+        \begin{example}
+        {
+            Conduct a goodness of fit test between the observed values below and the distribution $Po(2)$
+
+            \begin{center}
+            \begin{tabular}{l|c|c|c|c|c|c|c}
+            $x$ & 0 & 1 & 2 & 3 & 4 & 5 & $>5$ \\
+            \hline
+            $O_i$ & 12 & 23 & 24 & 24 & 12 & 5 & 0 \\
+            \end{tabular}
+            \end{center}
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0$: The observations \textit{can} be modeled by a $Po(2)$ distribution
+
+        $H_1$: The observations \textit{cannot} be modeled by a $Po(2)$ distribution
+        \end{step}
+
+        \begin{step}{Test statistic}
+        Work out $\chi^2$ using expected values from the $Po(2)$ distribution:
+
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c|c|c|c}
+        $x$ & 0 & 1 & 2 & 3 & 4 & 5 & $>5$ \\
+        \hline
+        $E_i$ & 13.53 & 27.07 & 27.07 & 18.04 & 9.02 & 3.61 & 1.66 \\
+        \end{tabular}
+        \end{center}
+
+        Since $E_i < 5$ we combine the last 2 columns:
+
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c|c|c}
+        $x$ & 0 & 1 & 2 & 3 & 4 & $>4$ \\
+        \hline
+        $O_i$ & 12 & 23 & 24 & 24 & 12 & 5 \\
+        $E_i$ & 13.53 & 27.07 & 27.07 & 18.04 & 9.02 & 5.27 \\
+        $\displaystyle\frac{(O_i - E_i)^2}{E_i}$ & 0.173 & 0.612 & 0.348 & 1.969 & 0.985 & 0.0138 \\
+        \end{tabular}
+        \end{center}
+
+        $$
+        \sum{\frac{(O_i - E_i)^2}{E_i}} = 4.10
+        $$
+        \end{step}
+
+        \begin{step}{Critical value}
+        \begin{align*}
+        v &= 6 - 1 \\
+        &= 5
+        \end{align*}
+        $$
+        \chi^2_6(5\%) = 11.07...
+        $$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        Since $4.10 < 11.07$ accept $H_0$ and reject $H_1$, the distribution $Po(2)$ is a good model.
+        \end{step}
+        \end{example}
+
+    \subsubsection{Continuous distributions}
+        For a continuous random variable, it is not possible to list all the values it can take. Therefore we divide the data into classes.
+
+        \begin{example}
+        {
+            The table below shows the time intervals in seconds between successive white cars in free-flowing traffic on an open raod.
+Investigate using a 5\% signigiicance level, whether these times can be modelled by an exponential distribution.
+
+            \begin{center}
+            \begin{tabular}{l|c|c|c|c|c|c}
+            Time & 0 - & 20 - & 40 - & 60 - & 90 - & 120 - 180 \\
+            Frequency & 41 & 19 & 16 & 13 & 9 & 2 \\
+            \end{tabular}
+            \end{center}
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0$: The observations \textit{can} be modeled by a exponential distribution
+
+        $H_1$: The observations \textit{cannot} be modeled by a exponential distribution
+        \end{step}
+
+        \begin{step}{Find an estimate for the parameter $\lambda$}
+            \begin{align*}
+            \bar{x} &= \frac{10 + 30 + 50 + 75 + 105 + 150}{6}\\
+                    &= 40
+            \end{align*}
+            The mean of an exponential distribution is $\displaystyle\frac{1}{\lambda}$ so the parameter $\lambda$ is $\displaystyle\frac{1}{40}$
+        \end{step}
+
+        \begin{step}{Find the probability that a value appears in each class}
+            The cumulative probabilty density function for this exponential distribution is
+            $$
+            F(x) = 1 - e^{-\lambda x}
+            $$
+            \begin{align*}
+            P(20) &= 0.3935 \\
+            P(40) - P(20) &= 0.2386 \\
+            P(60) - P(40) &= 0.1448 \\
+            P(90) - P(60) &= 0.1177 \\
+            P(120) - P(90) &= 0.0556 \\
+            P(180) - P(120) &= 0.0387 \\
+            P(\infty) - P(120) &= 0.0111 \\
+            \end{align*}
+        \end{step}
+
+        \begin{step}{Find the expected values for each class}
+        \begin{center}
+        \begin{tabular}{c|c|c|c}
+        Class & $O_i$ & Probability & $E_i$ \\
+        \hline
+        0- & 41 & 0.3935 & 39.35 \\
+        20- & 19 & 0.2386 & 23.86 \\
+        40- & 16 & 0.1448 & 14.48 \\
+        60- & 13 & 0.1177 & 11.77 \\
+        90- & 9 & 0.0556 & 5.56 \\
+        90- & 2 & 0.0387 & 3.87 \\
+        90- & 0 & 0.0556 & 1.11 \\
+        \end{tabular}
+        \end{center}
+
+        Combine last to columns so $E > 5$
+        \begin{center}
+        \begin{tabular}{c|c|c}
+        Class & $O_i$ & $E_i$ \\
+        \hline
+        0- & 41  & 39.35 \\
+        20- & 19 & 23.86 \\
+        40- & 16 & 14.48 \\
+        60- & 13 & 11.77 \\
+        90- & 9  & 5.56 \\
+        90- & 2  & 4.98 \\
+        \end{tabular}
+        \end{center}
+        Note that 4.98 < 5 however combining again would decrease the degree of freedom hence increasing the chance of a type 2 error.
+        \end{step}
+
+        \begin{step}{Test statistic}
+        $$
+        \sum{\frac{(O_i - E_i)^2}{E_i}} = 5.26
+        $$
+        \end{step}
+
+        \begin{step}{Critical value}
+        \begin{align*}
+        v &= 6 - 2 \\
+        &= 4
+        \end{align*}
+        $$
+        \chi^2_4(5\%) = 9.488...
+        $$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        Since $5.26 < 9.488$ accept $H_0$ and reject $H_1$, the exponential distribution with parameter $\lambda = \displaystyle\frac{1}{40}$ is a good model.
+        \end{step}
+
+        \end{example}

### 2017-04-11 src/stats6.tex
Created SS6 chapter structure.

\chapter*{Statistics 6}

\newpage
\section{Experimental Design}

\newpage
\section{Analysis of Variance}

\newpage
\section{Statistical Process Control}

\newpage
\section{Acceptance Sampling}

### 2017-04-12 src/stats6.tex
Added "Experimental design" to SS6.

@@
 
+    \newpage
+    \subsection{Experimental Design}
+        \paragraph{Experimental error}
+        Experimental error is the effect of other factors other that those controlled by the experimenter.
+The standard deviation is one way of measuring the experimental error, since its a measure of the spread of data.
+To minimize experimental error all factors that are not being investigated should be kept constant and a good experiment design should be used.
+
+        \paragraph{Experimental design}
+        The simplest experimental design is a paired comparision were experimental error is reduced bu applying the same treatment to the same subjects or in the same condition i.e.
+testing two diffrent diets on identical twins thus reducing experimental error due to phcological and biological effects.
+
+        \paragraph{Randomisation}
+        The purpose of randomisation is to reduce bias.
+Were possible the experimenter should remove all likely sources of bias by using random sampling to select an unbiased sample and random processes whenever their is a choise such as which twin is assigned which diet.
+
+        \paragraph{Blocking}
+        Completly randomised design is the process of randomly assigning subjects to treatments and uses one-factor analysis of varience.
+Randomised block design is the process of spliting subjects in to simular subgroups then randomly assigning those in a subgroups to a treatment.
+This design is analysed using two factor analysis of variance.
+Randomised block design is used when another factor (a blocking factor) is thought to an effect on the results.
+
+        \begin{example}
+            {
+                In a comparision of the drying times of three different types of wood preservatives A, B and C, samples of each four diffrent woods are availible, W1, W2, W3 and W4.
+Three experimental designs are suggested.
+                \begin{center}
+                \begin{tabular}{ccc|ccc|ccc}
+                & Design 1 & & & Design 2 & & & Design 3 & \\
+                A & B & C & A & B & C & A & B & C \\
+                \hline
+                W1 & W2 & W3 & W1 & W2 & W1 & W1 & W2 & W3 \\
+                W1 & W2 & W3 & W2 & W3 & W3 & W3 & W4 & W2 \\
+                W1 & W2 & W3 & W3 & W1 & W2 & W4 & W1 & W4 \\
+                &    &    & W4 & W4 & W4 & W2 & W3 & W1 \\
+                &    &    & W1 & W2 & W3 &    &    &    \\
+                \end{tabular}
+                \end{center}
+            }
+
+            \begin{step}{State two dissadvantages of design 1}
+            \begin{itemize}
+            \item Each type of preservative is used on only on type of wood thus it would not be possible to tell if any diffrence found was due to diffrent wood or the preservative (or both).
+            \item Wood 4 is not included in the test.
+            \end{itemize}
+            \end{step}
+
+            \begin{step}{Write down the name of design 3}
+            Randomised block design
+            \end{step}
+
+            \begin{step}{Name the tecnique used to analyse the results from design 3}
+            Two factor analysis of varience
+            \end{step}
+
+            \begin{step}{State one advantage of design 3 over design 2}
+            Design 2 uses some wood samples multiple times whereas design 3 uses all the samples only once.
+The comparision for mean drying times may be effected in design 2 by the cobination of wood samples used with each preservative.
+            \end{step}
+
+            \begin{step}{Explain how randomisation might be used in the context of design 1, 2 and 3}
+            \textbf{Design 1} Allocation of wood to treatment should be carried out by a random process, ensuring the results would not be rigged by selecting a wood for the mosted favored preservative.
+
+            \textbf{Design 2} A random process would decide which wood was W1 etc, thus the wood used twice would be selected randomly.
+
+            \textbf{Design 3} Less likely to be effected by bias but the order of carrying out treatments should ideally be determined by a random process. 
+            \end{step}
+        \end{example}
+
+        \paragraph{Control group}
+        If a new treatment is applied to an exprimental group, a control group, which recieves no treatment is needed to measure the effects of the new treatment.
+This control group should as simular as possible to the exprimental group in order to minimize bias.
+
+        \paragraph{Blind trials}
+        The plecbo effect is a well known effect were patients will improve by taking a plecbo drug, a treatment which contains no active ingredient.
+To show that a new drug is effective many more patients must show improvment than those taking the placbo pill.
+In a blind trial subjects do not know if they are receiving the treatment or the placbo which help reduce bias from the plecbo effect.
+
+        \paragraph{Double blind trials} Doctors that know a patient is on the active drug may expect patients fare better which may be transmitted onto the patient.
+A double blind trial is were the subject and the person administrating the treatment knows wether a placebo or active drug was given.
+
+        \paragraph{Triple blind trials} To prevent the bias in the statistical analysis it has been suggested that the statician should also not know which patients took the drug which would be a triple blind trial.
+
+        \begin{example}
+        {
+            To measure the effectiveness of a drug to relive breathlessness 12 subjects, all susceptible to breathlessness, were admisnistred the drug after one attack of breathlessness and the placebo after a seperate attack.
+One hour after the attacks an index of breathlessness was obtained for each subject, with the following results
+
+            \begin{center}
+            \begin{tabular}{c|c|c}
+            Subject & Drug & Placebo \\
+            \hline
+            1  & 28 & 32 \\
+            2  & 31 & 33 \\
+            3  & 17 & 23 \\
+            4  & 18 & 26 \\
+            5  & 31 & 34 \\
+            6  & 12 & 17 \\
+            7  & 33 & 30 \\
+            8  & 18 & 19 \\
+            9  & 25 & 23 \\
+            10 & 19 & 21 \\
+            11 & 17 & 24 \\
+            12 & 16 & 49 \\
+            \end{tabular}
+            \end{center}
+        }
+
+        \begin{step}{Explain the role that randomisation could play in carrying out the experiment}
+        \begin{itemize}
+        \item While it would be impossible to obtain a random sample of all sufferers of breathlessness, a random sample of thoses that are suitable and availible could be desided by a random process.
+        \item The order in which each patient takes the placbo and drug should be decided by a random process since the effect of the drug may be diffrent after the placbo (or visa-vera).
+        \item Alternativly a random selected group of 6 patients would take the placbo first, and the other group would take the drug first.
+        \end{itemize}
+        \end{step}
+
+        \begin{step}{Explain the meaning of a blind and of a double blind trial in the context of this expriment}
+        In a blind trial the subjects would not know weather they are taking the drug or placbo.
+In a double blind trial, the doctors that are administrating the drug or placbo and accessing the breathlessness index would also not know weather the subject was taking the drug or placbo.
+        \end{step}
+
+        \begin{step}{Making no assumption regarding the distribution of the data, investigate the claim (using 5\% significance level) that the drug significantly reduces the breathlessness index.}
+        \end{step}
+        $H_0: \text{Population median diffrence} = 0$\\
+        $H_1: \text{Population median diffrence} \ne 0$
+
+        \begin{center}
+        \begin{tabular}{c|c|c|c}
+        Subject & Drug & Placebo & Drug - Placebo \\
+        \hline
+        1  & 28 & 32 & $-$ \\
+        2  & 31 & 33 & $-$ \\
+        3  & 17 & 23 & $-$ \\
+        4  & 18 & 26 & $-$ \\
+        5  & 31 & 34 & $-$ \\
+        6  & 12 & 17 & $-$ \\
+        7  & 33 & 30 & $+$ \\
+        8  & 18 & 19 & $-$ \\
+        9  & 25 & 23 & $+$ \\
+        10 & 19 & 21 & $-$ \\
+        11 & 17 & 24 & $-$ \\
+        12 & 16 & 49 & $-$ \\
+        \end{tabular}
+        \end{center}
+
+        $$
+        10^- / 2^+
+        $$
+
+        \begin{align*}
+        X &\sim B(12, 0.5) \\
+        P(X \leq 2) &= P(X < 3) \\
+        &= 0.0193
+        \end{align*}
+
+        Since $0.0193 < 0.05$ reject $H_0$ at the 5\% significance level the drug reduceses breathlessness.
+
+        \end{example}
+
+    \newpage
+    \subsection{Analysis of paired comparisions}
+        Assuming their are two normal populations from which paired samples of size $n$ are taken, with means $\mu_1$ and $\mu_2$, then the diffrence between the pairs will also be normal distributed.
+Thus a test for $\mu_1 = \mu_2$ is equivilent to $\mu_d = 0$.
+
+        $$
+        \bar{D} \sim N\left(\mu_d, \frac{\sigma_d^2}{n}\right)
+        $$
+
+        $$
+        Z = \dfrac{\bar{D} - \mu_d}{\dfrac{\sigma_d}{\sqrt{n}}} \sim N(0, 1)
+        $$
+        Let $\bar{D}$ and $S_d^2$ denote the mean and variance of the a sample of n diffrences. Thus
+        $$
+        \frac{\bar{D} - \mu_d}{\dfrac{S_d}{\sqrt{n}}} \sim t_{n-1}
+        $$
+
+        \begin{example}
+        {
+            A school mathamatics teachers wants to test the effect of a new educational computer package.
+She selects pairs of students of equal ability and randomly selects one from each pair to join the experimental group (with the others making the control group).
+The results of a later assesment are as follows:
+
+            \begin{center}
+            \begin{tabular}{c|c|c}
+            Pair & Control & Experimental \\
+            \hline
+            1  & 72 & 75 \\
+            2  & 82 & 79 \\
+            3  & 93 & 84 \\
+            4  & 65 & 71 \\
+            5  & 76 & 82 \\
+            6  & 89 & 91 \\
+            7  & 81 & 85 \\
+            8  & 58 & 68 \\
+            9  & 95 & 90 \\
+            10 & 91 & 92 \\
+            \end{tabular}
+            \end{center}
+
+            Assuming the diffrence in marks are normally distributed, investigate the claim that the computer package inproves students understanding using a 5\% significance level.
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0: \mu_d = 0$\\
+        $H_0: \mu_d > 0$
+        \end{step}
+
+        \begin{step}{Test statistic}
+        \begin{align*}
+        \bar{d} &= 1.5\\
+        s_d &= 5.720\\
+        t &= \frac{1.5 - 0}{\dfrac{5.720}{\sqrt{10}}}\\
+        &= 0.8293
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Critical value}
+        $$t_{10 - 1}(0.95) = 1.833$$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $0.8293 < 1.833$ so accept $H_0$, their is no evidence at the 5\% significance level that the computer package improves the students results.
+        \end{step}
+
+        \end{example}
 \newpage

### 2017-04-12 src/stats6.tex
Added "Analysis of Variance" to SS6.

@@
         \end{example}
+
 \newpage
 \section{Analysis of Variance}
+ANOVA tests are an extension of F-Tests for tests with more than 2 populations. When dealing with anova tests their are two new terms used:
+\paragraph{Factor} A characteristic under consideration, thought to influence the measured observations.
+\paragraph{Level} A value of the factor.\\
+\\
+Consider the following test:
+A car magazine wishes to compare the average petrol consumption of three simular models of car and has availible six vehicles of each model. \\
+\\
+This test has 1 factor (model of car) at 3 levels (3 diffrent models).
+
+    \newpage
+    \subsection{One-way ANOVA}
+        A one-way ANOVA test is used when their is only 1 factor. The follwing notation and formule are used in the calculation of a one-way ANOVA test:
+        \bgroup
+        \def\arraystretch{2}
+        \begin{center}
+        \begin{tabular}{l|c|c}
+        Description & Symbol & Value \\
+        \hline
+        Number of samples/levels & k & \\
+        Number of observations in $i^{th}$ sample & $n_i$ & $i = 1, 2, ..., k$ \\
+        Total number of observations & $n$ & $\displaystyle\sum_i{n_i}$ \\
+        Observation $j$ in $i^{th}$ sample & $x_{ij}$ & $j = 1, 2, ..., n_i$ \\
+        Sum of $n_i$ observations in the $i^{th}$ sample & $T_i$ & $\displaystyle\sum_j{x_{ij}}$ \\
+        Sum of all n observations & $T$ & $\displaystyle\sum_i{T_i}$ \\
+        \hline
+        Total sum of squares & $SS_T$ & $\displaystyle\sum_i{\displaystyle\sum_j{x_{ij}^2}} - \frac{T^2}{n}$ \\
+        Between samples sum of squares & $SS_B$ & $\displaystyle\sum_i{\frac{T_i}{n_i}} - \frac{T^2}{n}$ \\
+        Within samples sum of squares & $SS_W$ & $SS_T - SS_B$ \\
+        \hline
+        Between samples mean squares & $MS_B$ & $\displaystyle\frac{SS_B}{k - 1}$ \\
+        Within samples mean squares& $MS_W$ & $\displaystyle\frac{SS_W}{n - k}$ \\
+        Total mean squares & $MS_T$ & $\displaystyle\frac{SS_T}{n - 1}$ \\
+        \end{tabular}
+        \end{center}
+        \egroup
+
+        \noindent
+        Its convient to lay out an ANOVA test in a table such as:
+        \bgroup
+        \def\arraystretch{2}
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c}
+        Source of variation & Sum of squares    & Degrees of freedom    & Mean square   & F-ratio               \\
+        \hline
+        Between samples     & $SS_B$            & $k-1$                 & $MS_B$        & $\dfrac{MS_B}{MS_W}$  \\
+        Within samples      & $SS_W$            & $n-k$                 & $MS_W$        &                       \\
+        \hline
+        Total               & $SS_T$            & $n-1$                 &               &                       \\
+        \end{tabular}
+        \end{center}
+        \egroup
+        The F-ratio forms the test statistic and the F-distribution is used to find the critical value.
+Note that $MS_B$ will always be larger than $MS_W$ so we always select the upper tail as the critical value.
+
+        \begin{example}
+        {   
+            In a comparision of the cleaning action of four detergents, pieces of cloth were measured for whiteness after being washed by a particular detergent. The results were as follows:
+            \begin{center}
+            \begin{tabular}{c|c|c|c}
+            A  & B  & C  & D  \\
+            \hline
+            77 & 74 & 73 & 76 \\
+            81 & 66 & 78 & 85 \\
+            61 & 58 & 57 & 77 \\
+            76 &    & 69 & 64 \\
+            69 &    & 63 &    \\
+            \end{tabular}
+            \end{center}
+            Assuming the whiteness readings are normally distributed, with a common variance, test hypothesis that their is no diffrence between the four brands whiteness using a 5\% significance level.
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0$: no diffrence in the mean whiteness\\
+        $H_1$: a diffrence in the mean whiteness\\
+        \end{step}
+
+        \begin{step}{Work out the sum of squares}
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c|c}
+                & A     & B     & C     & D     & Total \\
+        \hline
+        $n_i$   & 5     & 3     & 5     & 4     & 17    \\
+        $T_i$   & 364   & 198   & 340   & 302   & 1204  \\
+        \end{tabular}
+        \end{center}
+
+        \begin{align*}
+        SS_T &= 77^2 + 81^2 + 61^2 + ... + 64^2 - \frac{1204^2}{17}\\
+        &= 86362 - 85271.529...\\
+        &= 1090.47\\
+        \\
+        SS_B &= (\frac{364^2}{5} + \frac{198^2}{3} + \frac{340^2}{5} + \frac{302^2}{4}) - \frac{1204^2}{17}\\
+        &= 85488.2 - 85271.529...\\
+        &= 216.67\\
+        \\
+        SS_W &= 1090.47 - 216.67\\
+        &= 873.80
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Complete the ANOVA table}
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c}
+        Source of variation & Sum of squares    & Degrees of freedom    & Mean square   & F-ratio   \\
+        \hline
+        Between samples     & $216.67$          & $3$                   & $72.22$       & $1.07$    \\
+        Within samples      & $873.80$          & $13$                  & $67.22$       &           \\
+        \hline
+        Total               & $1090.47$         & $16$                  &               &           \\
+        \end{tabular}
+        \end{center}
+        \end{step}
+
+        \begin{step}{Critical value}
+        $$
+        F_{(3, 13)}(0.95) = 3.411 
+        $$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $1.07 < 3.411$ so accept $H_0$, their is no evidence at the 5\% significance level to suggest that their is any diffrence between the four brands.
+        \end{step}
+        \end{example}
+
+    \newpage
+    \subsection{Two-way ANOVA}
+        A two-way ANOVA is an extension to the one-way ANOVA test that takes into account a second factor.
+The second factor is the blocking factor (since this test requires a radomised block design) since it places the samples into blocks.
+        $$
+        x_{ij} - \mu_{ij} = \epsilon_{ij} \sim N(0, \sigma^2)
+        $$
+        where
+        $$
+        \mu_{ij} = \mu + \alpha_i + \beta_j
+        $$
+        where $\mu$ is the overall mean, $\alpha_i$ is the mean effect of the $i^{th}$ level of the row factor relative to $\mu$, $\beta_j$ is the mean effect of the $j^{th}$ level of the column factor relative to $\mu$ and $\epsilon_{ij}$ is the inherent random variation.\\
+        \\
+        The follwing notation and formule are used in the calculation of a two-way ANOVA test:
+        \bgroup
+        \def\arraystretch{2}
+        \begin{center}
+        \begin{tabular}{l|c|c}
+        Description & Symbol & Value \\
+        \hline
+        Number of levels of row factor & $m$ & \\
+        Number of levels of column factor & $n$ & \\
+        Total number of observations & $mn$ & \\
+        Observation in $ij^{th}$ cell & $x_{ij}$ & $i = 1, 2, ..., m$ \\
+        & & $j = 1, 2, ..., n$  \\
+        Sum of $n$ observations in the $i^{th}$ row & $R_i$ & $\sum_j{x_{ij}}$ \\
+        Sum of $m$ observations in the $j^{th}$ row & $C_j$ & $\sum_i{x_{ij}}$ \\
+        Sum of all $mn$ observations & $T$ & $\sum_i{\sum_j{x_{ij}}} = \sum_i{R_i} = \sum_j{R_j}$ \\
+        \hline
+        Total sum of squares & $SS_T$ & $\displaystyle\sum_i{\displaystyle\sum_j{x_{ij}^2}} - \frac{T^2}{mn}$ \\
+        Between rows sum of squares & $SS_R$ & $\displaystyle\sum_i{\frac{R_i^2}{n}} - \frac{T^2}{mn}$ \\
+        Between column sum of squares & $SS_C$ & $\displaystyle\sum_i{\frac{C_j^2}{m}} - \frac{T^2}{mn}$ \\
+        Error (residual) sum of squares & $SS_E$ & $SS_T - SS_R - SS_C$ \\
+        \hline
+        Between row mean squares & $MS_R$ & $\displaystyle\frac{SS_R}{n - 1}$ \\
+        Between column mean squares & $MS_C$ & $\displaystyle\frac{SS_C}{m - 1}$ \\
+        Error (residual) mean squares & $MS_E$ & $\displaystyle\frac{SS_E}{mn - 1}$ \\
+        \end{tabular}
+        \end{center}
+        \egroup
+        \noindent
+
+        Its convient to lay out a two-way ANOVA test in a table such as:
+        \bgroup
+        \def\arraystretch{2}
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c}
+        Source of variation & Sum of squares    & Degrees of freedom    & Mean square   & F-ratio               \\
+        \hline
+        Between rows        & $SS_R$            & $m-1$                 & $MS_R$        & $\dfrac{MS_R}{MS_E}$  \\
+        Between columns     & $SS_C$            & $n-1$                 & $MS_C$        & $\dfrac{MS_C}{MS_E}$  \\
+        Error (residual)    & $SS_E$            & $(m-1)(n-1)$          & $MS_E$        &                       \\
+        \hline
+        Total               & $SS_T$            & $mn-1$                &               &                       \\
+        \end{tabular}
+        \end{center}
+        \egroup
+
+        \begin{example}
+        {
+            Test at a 1\% significance level, that their is no diffrence between the four compilers and was the use of the program as a blocking factor worthwhile.
+
+            \begin{center}
+            \begin{tabular}{c|c|c|c|c}
+            & \multicolumn{4}{c}{Compiler} \\
+            Program & 1         & 2         & 3         & 4         \\
+            \hline
+            A       & $29.21$   & $28.25$   & $28.20$   & $28.62$   \\
+            B       & $26.18$   & $26.02$   & $26.22$   & $25.56$   \\
+            C       & $30.91$   & $30.18$   & $30.52$   & $30.09$   \\
+            D       & $25.14$   & $25.26$   & $25.20$   & $25.02$   \\
+            E       & $26.16$   & $25.14$   & $25.26$   & $25.46$   \\
+            \end{tabular}
+            \end{center}
+        }
+
+        \begin{step}{Hypothesis}
+        $H_0$: No effect on compilation times due to compilers\\
+        $H_1$: An effect on compilation times due to compilers\\
+        \end{step}
+
+        \begin{step}{Find the row and column totals}
+        \begin{center}
+        \begin{tabular}{c|cccc|c|c}
+                            & \multicolumn{4}{c}{Compiler}                  & Row totals    &           \\
+        Program             & 1         & 2         & 3         & 4         & $T_i$         & $n_i$     \\
+        \hline          
+        A                   & $29.21$   & $28.25$   & $28.20$   & $28.62$   & $114.28$      & 4         \\
+        B                   & $26.18$   & $26.02$   & $26.22$   & $25.56$   & $103.98$      & 4         \\
+        C                   & $30.91$   & $30.18$   & $30.52$   & $30.09$   & $121.70$      & 4         \\
+        D                   & $25.14$   & $25.26$   & $25.20$   & $25.02$   & $100.62$      & 4         \\
+        E                   & $26.16$   & $25.14$   & $25.26$   & $25.46$   & $102.02$      & 4         \\
+        \hline
+        Column totals $T_j$ & $137.60$  & $134.85$  & $135.40$  & $134.75$  & $T = 542.60$  &           \\
+        \hline
+        $n_j$               & $5$       & $5$       & $5$       & $5$       &               & $mn = 20$ \\
+        \end{tabular}
+        \end{center}
+        \end{step}
+
+        \begin{step}{Find the sum of squares}
+        \begin{align*}
+        SS_T &= (29.21^2 + 28.25^2 + ... + 25.46^2) - \frac{542.60^2}{20} \\
+        &= 14805.7768 - 14720.738\\
+        &= 85.0388\\
+        \\
+        SS_R &= \left(\frac{114.28^2}{4} + \frac{103.98^2}{4} + \frac{121.70^2}{4} + \frac{100.62^2}{4} + \frac{102.02^2}{4}\right) - \frac{542.60^2}{20}\\
+        &= 83.0404\\
+        \\
+        SS_C &= \left(\frac{137.60^2}{5} + \frac{134.85^2}{5} + \frac{135.40^2}{5} + \frac{134.75^2}{5}\right) - \frac{542.60^2}{20}\\
+        &= 1.063\\
+        \\
+        SS_E &= 85.0388 - 83.0404 - 1.063\\
+        &= 0.9354
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Complete the ANOVA table}
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c}
+        Source of variation & Sum of squares    & Degrees of freedom    & Mean square   & F-ratio       \\
+        \hline
+        Between rows        & $83.0404$         & $4$                   & $20.7601$     & $266.33$      \\
+        Between columns     & $1.063$           & $3$                   & $0.3543$      & $4.55$        \\
+        Error (residual)    & $0.9354$          & $12$                  & $0.07795$     &               \\
+        \hline
+        Total               & $85.0388$         & $11$                  &               &               \\
+        \end{tabular}
+        \end{center}
+        \end{step}
+
+        \begin{step}{Critical value}
+        $$
+        F_{(3, 12)}(0.99) = 5.953
+        $$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        $4.55 < 5.953$ so accept $H_0$, thus their is no evidence at the 1\% significance level to suggest a diffrence in compilation times between the four compilers.
+        \end{step}
+
+        \begin{step}{Was the blocking factor worthwhile?}
+        \begin{itemize}
+        \item The variation between programs accounted for $\dfrac{83.0404}{85.0388} \times 100 = 97.65\%$ of the variation (which most would have gone into $SS_E$ had the blocking factor not been used).
+        \item $F_R = 266.33$ which would have been significant at any level.
+        \end{itemize}
+        \end{step}
+
+        \end{example}
+
+    \newpage
+    \subsection{Latin square designs}
+        Consider a randomised block design in which:
+        $$ \text{number of blocks} = \text{number of treatments} = n $$
+        This is equivilent to a two-way ANOVA in which:
+        $$ \text{number of rows} = \text{number of columns} = n $$
+        So the total number of observations is $n^2$.\\
+        \\
+        It is possible to intoduce a second blocking factor (with $n$ levels) without increasing the number of observations to $n^3$ with the following model:
+        $$
+        x_{ij(k)} - \mu_{ij(k)} = \epsilon_{ij(k)} \sim N(0, \sigma^2)
+        $$
+        where 
+        \begin{itemize}
+        \item $x_{ij(k)}$ is the observation in the $i^{th}$ row, $j^{th}$ column with the letter $k$
+        \item $\mu$ is the overall mean
+        \item $\alpha_i$ is the mean effect of the $i^{th}$ level of row factor relative to $\mu$
+        \item $\beta_j$ is the mean effect of the $j^{th}$ level of column factor relative to $\mu$
+        \item $\gamma_k$ is the mean effect of the $k^{th}$ level of letter factor relative to $\mu$
+        \item $\epsilon_{ij(k)}$ is the inherent random variation 
+        \end{itemize}
+
+        \begin{example}
+        {
+            In a comparision of three types of fuels (A, B and C), three makes of cars (1, 2 and 3), were used.
+Each car was driven at a speed of $50 km h^-1$, $75 km h^-1$ and $100 km h^-1$ around a track.
+The distrance travelled from depleting a full tank was recorded.
+
+            \begin{center}
+            \begin{tabular}{c|c|c|c}
+            Car & Speed & Fuel  & Distance  \\
+            1   & 50    & B     & 347.8     \\
+            1   & 75    & C     & 331.4     \\
+            1   & 100   & A     & 275.8     \\
+            2   & 50    & A     & 365.4     \\
+            2   & 100   & B     & 339.8     \\
+            2   & 150   & C     & 298.9     \\
+            3   & 50    & C     & 318.9     \\
+            3   & 100   & A     & 302.2     \\
+            3   & 150   & B     & 271.8     \\
+            \end{tabular}
+            \end{center}
+
+            Investigate the diffrence between the three petrols.
+        }
+
+        \begin{step}{Rearrange the data into a latin square}
+        \begin{center}
+        \begin{tabular}{c|ccc|c|c}
+                        & \multicolumn{3}{c}{Speed}         &           &           \\
+        Car             & 50        & 75        & 100       & Row Total & $n_i$     \\
+        \hline
+        1               & B $347.8$ & C $331.4$ & A $275.8$ & $955.0$       & 3     \\
+        2               & A $365.4$ & B $339.8$ & C $298.9$ & $1004.1$      & 3     \\
+        3               & C $318.9$ & A $302.2$ & B $271.8$ & $892.9$       & 3     \\
+        \hline
+        Column total    & $1032.1$  & $973.4$   & $846.5$   & $T = 2852.0$  &       \\
+        \hline
+        $n_j$           & 3         & 3         & 3         &               & n = 9 \\
+        \end{tabular}
+        \end{center}
+
+        \begin{align*}
+        L_A &= 943.4\\
+        L_B &= 959.4\\
+        L_C &= 949.2
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Find the sum of squares}
+        \begin{align*}
+        SS_T &= (347.8^2 + 331.4^2 + ... + 271.8^2) - \frac{2852.0^2}{9}\\
+        &= 912076.14 - 903767.11\\
+        &= 8309.03\\
+        \\
+        SS_R &= (\frac{955.0^2}{3} + \frac{1004.1^2}{3} + \frac{892.9^2}{3}) - 903767.11\\
+        &= 2070.30\\
+        \\
+        SS_C &= (\frac{1032.1^2}{3} + \frac{973.4^2}{3} + \frac{846.5^2}{3}) - 903767.11\\
+        &= 5999.63\\
+        \\
+        SS_L &= (\frac{943.4^2}{3} + \frac{959.4^2}{3} + \frac{949.2^2}{3}) - 903767.11\\
+        &= 43.74\\
+        \\
+        SS_E &= 8309.03 - 2070.30 - 5999.63 - 43.74
+        &= 195.36
+        \end{align*}
+        \end{step}
+
+        \begin{step}{Contruct ANOVA table}
+        \begin{center}
+        \begin{tabular}{l|c|c|c|c}
+        Source of variation & Sum of squares    & Degrees of freedom    & Mean square   & F-ratio       \\
+        \hline
+        Between cars        & $2070.30$         & $2$                   & $1035.15$     & $10.60$       \\
+        Between speeds      & $5999.63$         & $2$                   & $2999.81$     & $30.71$       \\
+        Between brands      & $43.74$           & $2$                   & $21.87$       & $0.22$        \\
+        Residual            & $0.9354$          & $2$                   & $97.68$       &               \\
+        \hline
+        Total               & $8309.03$         & $8$                   &               &               \\
+        \end{tabular}
+        \end{center}
+        \end{step}
+
+        \begin{step}{Critical value}
+        $$
+        F_{(2, 2)}(0.95) = 19.00
+        $$
+        \end{step}
+
+        \begin{step}{Conclusion}
+        \begin{itemize}
+        \item $0.22 < 19.00$ so their is no evidence that fuel consumption is dependent on the brand of fuel
+        \item $30.71 > 19.00$ so the speed of cars proved to be signigicant
+        \item $10.60 < 19.00$ so the car was not suggnificant, however it did reduce the residual sum of squares 
+        \end{itemize}
+        \end{step}
+
+        \end{example}
 

### 2017-04-14 src/stats6.tex
Added "Statistical Process Control" to SS6.

@@
     \newpage
+    \subsection{Control charts}
+        \paragraph{Statistical process control}
+        Detect and signal when a production lines item-to-item varibility has increased or the process mean has moved form the target.
+
+    \subsubsection{Control of means}
+        The most common method of statistical process control is to take regular samples and plot them on a control chart as shown in the following diagram:\\
+        \includegraphics[scale=0.125]{controlchart}
+
+        \begin{center}
+        \begin{tabularx}{\textwidth}{l|X}
+        Event & Action \\
+        \hline
+        Sample mean is between warning limits & No action \\
+        Sample mean between warning and action limits & Take another sample immediently, if the new sample mean is outside warning limits take action \\
+        Sample mean outside action limits & Take action \\
+        \end{tabularx}
+        \end{center}
+
+        \noindent
+        To set the limits of a control chart a large sample (when the process is nominal) is taken, known as a proccess capability study.
+Then find the standard deviation $\sigma$, the size of the continous samples $n$ and the critical values (normaly 95\% and 99.8\%).
+With this data the confidence itervals for the warning and action limits can be found.
+
+        \begin{example}
+        {
+            A company manufactures panels for use in making baths.
+The panels have a target width of 700mm.
+When a production is satisfactory the widths are normally distributed with a standard deviation of 2mm.
+Production is to be controlled by taking a regualar sample of 5 panels.
+        }
+
+        \begin{step}{Calculate the warning and action limits for a control chart}
+        \begin{align*}
+        \text{warning limits} &= 700 \pm 1.96 \times \frac{2}{\sqrt{5}}\\
+        &= (698.247, 701.753)\\
+        &= (698.2, 701.8)
+        \end{align*}
+        \begin{align*}
+        \text{action limits} &= 700 \pm 3.09 \times \frac{2}{\sqrt{5}}\\
+        &= (697.236, 702.764) \\
+        &= (697.2, 702.8)
+        \end{align*}
+        \end{step}
+
+        \begin{step}{What action would you take for the following sample $701.2, 698.2, 704.4, 699.4, 695.5$}
+        $\bar{x}$ = 699.7\\
+        Mean is within warning limits so no action taken.
+        \end{step}
+
+        \begin{step}{What action would you take for the following sample $700.2, 697.5, 695.1, 696.0, 698.9$}
+        $\bar{x}$ = 697.5\\
+        Mean is between warning limit and action limit so another sample should be taken immediently, it it is outside of the warning limits action should be taken.
+        \end{step}
+
+        \begin{step}{What action would you take for the following sample $699.5, 707.1, 704.9, 703.9, 706.4$}
+        $\bar{x}$ = 704.4\\
+        Mean is outside action limits so take action immediently.
+        \end{step}
+
+        \end{example}
+
+        \paragraph{Target values} Sometimes the target value is not defined/unattainable, it this case a large samples sample mean could be used.
+In the case were a warning that a process is doint better than expected (i.e.
+the strength of a iron casting is stronger than expected) an invistigation should be ran to see how the improvement can be maintained.
+
+    \subsubsection{Control of varibility}
+        Within a sample their may be high varibility but a normal looking mean, to detect these events we can use a chart for range or a chart for standard deviation.
+To get the upper and lower limits for these charts multiply the short-term standard deviation by the appropiate value of D/E on table 12 (Control charts for varibility).
+Standard deviation charts have a better chance to detect a change in varibility with the same rate of false positives as a range chart, however the range is simpler to calculate.
+
+        \begin{example}
+        {
+            The same company (as in the previous example) wants to control varibility.
+        }
+
+        \begin{step}{Calculate upper and lower warning limits for a standard deviation chart}
+        \begin{itemize}
+        \item Upper action limit $2.15 \times 2 = 4.30$
+        \item Upper warning limit $1.67 \times 2 = 3.34$
+        \item Lower warning limit $0.35 \times 2 = 0.70$
+        \item Lower action limit $0.15 \times 2 = 0.30$
+        \end{itemize}
+        \end{step}
+
+        \begin{step}{What action would you take for the following sample $701.2, 698.2, 704.4, 699.4, 695.5$}
+        $\sigma = 3.328$\\
+        Standard deviation is within warning limits so no action taken.
+        \end{step}
+
+        \begin{step}{What action would you take for the following sample $700.2, 697.5, 695.1, 696.0, 698.9$}
+        $\sigma = 2.077$\\
+        Standard deviation is within warning limits so no action taken.
+        \end{step}
+
+        \begin{step}{What action would you take for the following sample $699.5, 707.1, 704.9, 703.9, 706.4$}
+        $\sigma = 2.991$\\
+        Standard deviation is within warning limits so no action taken.
+        \end{step}
+
+        \end{example}
+
+    \subsubsection{Estimating short-test standard deviation}
+        Sometimes its not possible to collect a large sample to find the target value, instead we can estimate it from the range of several samples.
+Find the range of all the samples then find the mean range.
+Multiply this value by the suitable b value from table 12 to find an estimate for the standard deviation.
+Alternativly the b value can be applied to all the ranges and the mean of those values can be used.
+
+        \begin{example}
+        {
+            9 samples of size 5 from a production process have ranges of:
+            $$
+            2.3, 2.9, 1.8, 3.4, 2.0, 3.6, 2.9, 2.1, 1.1
+            $$
+            Estimate the standard deviation of the process
+        }
+
+        \begin{step}{}
+            \begin{align*}
+            \text{mean range} &= \frac{2.3+2.9+1.8+3.4+2.0+3.6+2.9+2.1+1.1}{9}\\
+            &= \frac{16.5}{9}\\
+            &= 1.8333\\
+            \\
+            1.8333 \times 0.4299 &= 0.7881\\
+            &= 0.788\\
+            \end{align*}
+        \end{step}
+
+        \end{example}
+
+    \subsubsection{Tolerance limits}
+        Many processes have limits which the product \textit{must} conform to.
+For example drugs imagine a drug with a tollerance of 4.9 to 5.1 (for potency) with a target value of 5.0.
+Its esimated standard deviation was 0.063 which gives the following upper and lower limits.
+        \begin{align*}
+        z_1 &= \frac{4.9-5.0}{0.063}\\
+        &= -1.587\\
+        \\
+        z_2 &= \frac{5.1-5.0}{0.063}\\
+        &= 1.587
+        \end{align*}
+        which means $1 - 0.887 = 11.3\%$ of the drugs would be outside the tollerance limits.\\
+        \\
+        If a process's tollerance width exceeds 6 standard deviations the process can be said to meet the tolernaces consitently, given that the mean is kept on target.
+The drug had a tollerance width of $\displaystyle\frac{0.51 - 0.49}{0.063} = 3.2$ standard deviations.
+
+        \begin{example}
+        {
+            A biscuit factory produces cream crackers.
+Packets are sampled at hourly intervals and the weight, in grams, of the contents measured.
+The results below are from seven samples each of size 5, taken at a time when production was thought to be satisfactory.
+The target value is 210g.
+
+            \begin{center}
+            \begin{tabular}{c|c|c|c|c|c|c}
+            \multicolumn{7}{c}{Sample}\\
+            1         & 2         & 3         & 4         & 5         & 6         & 7         \\
+            \hline
+            $209.0$   & $214.5$   & $204.5$   & $217.5$   & $211.0$   & $224.0$   & $210.0$   \\
+            $211.0$   & $210.0$   & $209.0$   & $216.0$   & $198.0$   & $220.0$   & $208.0$   \\
+            $213.5$   & $212.0$   & $198.5$   & $217.0$   & $209.5$   & $224.0$   & $210.0$   \\
+            $205.0$   & $203.5$   & $203.0$   & $214.0$   & $213.5$   & $218.5$   & $220.0$   \\
+            $210.0$   & $214.0$   & $213.5$   & $209.5$   & $213.5$   & $210.0$   & $211.0$   \\
+            \end{tabular}
+            \end{center}
+        }
+
+        \begin{step}{Estimate the standard deviation of the process}
+        $$\text{ranges} = 8.5, 11, 15, 8, 15.5, 14, 12$$
+        $$\text{mean range} = 12$$
+        \begin{align*}
+        \sigma &= b \times 12 \\
+        &= 0.4299 \times 12 \\
+        &= 5.1588 \\
+        &= 5.16
+        \end{align*}
+        \end{step}
+
+        \begin{step}{The tollerance for this process is 200g to 220g, comment on the the ability to meet these tollerances}
+        The tollerance width is $\displaystyle\frac{220 - 200}{5.16} = 3.88$ standard deviations wide which is bellow the 6 standard deviations we would expect for a process to meet tollerances consistently.
+        \end{step}
+
+        \end{example}
+
+    \subsubsection{Control of proportion}
+        Control charts may also be used when items are classified as conforming/non-conforming (instead of by weight or height etc).
+Consider a production line were samples were taken when the production was thought to be satisfactory.
+With a sample size of 100, the following are the number of non-conforming items in each sample:
+        $$
+        10, 13, 12, 19, 8, 14, 17, 16, 10, 18, 9, 16
+        $$
+        The total number of non-conforming items are 162, so the estimate for the proportion of non-conforming items would be $\hat{p} = \dfrac{162}{1000} = 0.135$.
+Given $\hat{p}$ and sample size $n$, we can find the confidence interval for proportion hence we can find the action and warning limits
+
+        \begin{example}
+        {
+            To control a process producing components for washing machines 75 components are tested every hour and the number of non-conforming components are recorded.
+The numbers recorded when production was regarded as satisfactory were:
+            $$
+            6, 16, 12, 9, 14, 19, 7, 13
+            $$
+        }
+
+        \begin{step}{Find the warning and action limits for a control chart of proportion}
+        \begin{align*}
+        p &= \frac{6 + 16 + 12 + 9 + 14 + 19 + 7 + 13}{75 * 8}\\
+        &= \frac{96}{600}\\
+        &= 0.16
+        \end{align*}
+
+        \begin{align*}
+        \text{action limits} &= 0.16 \pm 3.09 \times \sqrt{\frac{0.16 \times (1 - 0.16)}{75}} \\ 
+        &= (0.0292, 0.2908)\\
+        \text{warning limits} &= 0.16 \pm 1.96 \times \sqrt{\frac{0.16 \times (1 - 0.16)}{75}} \\
+        &= (0.0770, 0.2430)
+        \end{align*}
+        \end{step}
+
+        \begin{step}{What action would be taken if the next sample was 27}
+        $\dfrac{27}{75} = 0.36$ which is outside the action limits so take action.
+        \end{step}
+
+        \begin{step}{What action would be taken if the next sample was 15}
+        $\dfrac{15}{75} = 0.2$ which is insied the warning limits so no action is taken.
+        \end{step}
+
+        \begin{step}{What action would be taken if the next sample was 3}
+        $\dfrac{3}{75} = 0.04$ is bellow the warning limit so investigate how this low rate of non-conforming items may be maintained.
+        \end{step}
+
+        \end{example}
+
+\newpage
 \section{Acceptance Sampling}

### 2017-04-14 src/stats6.tex
Converted control chart to tikz picture.

@@
         The most common method of statistical process control is to take regular samples and plot them on a control chart as shown in the following diagram:\\
-        \includegraphics[scale=0.125]{controlchart}
+
+        \newcommand{\xpoint}[2]{
+            \draw[line width=1] (#1-0.1,#2-0.1) -- (#1+0.1, #2+0.1);
+            \draw[line width=1] (#1-0.1,#2+0.1) -- (#1+0.1, #2-0.1);
+        }
+
+        \newcommand{\xpointn}[3]{
+            \draw[line width=1] (#1-0.1,#2-0.1) -- (#1+0.1, #2+0.1);
+            \draw[line width=1] (#1-0.1,#2+0.1) -- (#1+0.1, #2-0.1) node[right]{#3};
+        }
+
+        \begin{center}
+        \begin{tikzpicture}
+        \draw[->, line width=1] (1, 1) -- (1, 5) node[left]{
+            \begin{tabular}{r} Sample \\ mean \end{tabular}
+        };
+        \draw[line width=1] (1, 4.5) -- (10, 4.5) node[right]{Action limit};
+        \draw[line width=1] (1, 4) -- (10, 4) node[right]{Warning limit};
+        \draw[line width=1] (1, 3) -- (10, 3) node[right]{Target value};
+        \draw[line width=1] (1, 2) -- (10, 2) node[right]{Warning limit};
+        \draw[line width=1] (1, 1.5) -- (10, 1.5) node[right]{Action limit};
+        \xpoint{1.6}{3.2}
+        \xpoint{2.4}{2.5}
+        \xpoint{3.2}{3.4}
+        \xpoint{3.7}{3.7}
+        \xpoint{4.1}{3.2}
+        \xpoint{4.6}{2.2}
+        \xpoint{5.3}{3.1}
+        \xpoint{6.1}{3.4}
+        \xpoint{6.7}{2.8}
+
+        \xpointn{7.4}{3.8}{(1)}
+        \xpointn{7.7}{1.2}{(2)}
+        \xpointn{8.7}{4.25}{(3)}
+        \end{tikzpicture}
+        \end{center}
 
@@
         \hline
-        Sample mean is between warning limits & No action \\
-        Sample mean between warning and action limits & Take another sample immediently, if the new sample mean is outside warning limits take action \\
-        Sample mean outside action limits & Take action \\
+        (1) Sample mean is between warning limits & No action \\
+        (2) Sample mean outside action limits & Take action \\
+        (3) Sample mean between warning and action limits & Take another sample immediently, if the new sample mean is outside warning limits take action \\
         \end{tabularx}

### 2017-04-15 src/stats6.tex

@@
 
+    \newpage
     \subsubsection{Control of means}
@@
 
+    \newpage
     \subsubsection{Control of varibility}
@@
 
+    \newpage
     \subsubsection{Estimating short-test standard deviation}
@@
 
+    \newpage
     \subsubsection{Tolerance limits}
@@
 
+    \newpage
     \subsubsection{Control of proportion}
@@
 \section{Acceptance Sampling}
+
+    \paragraph{Acceptance Sampling}
+    The process of deciding if a large batch of simular items is acceptable by testing a small amount of them.
+
+    \paragraph{Acceptance sampling by attributes} 
+    Each item is classified as conforming or non-conforming, if the sample contains to many non-conforming items the batch is rejected.
+
+    \paragraph{Operating characteristic}
+    A graph of probabilty of acceptance against proportion of non-conforming in the batch. Providing the sample size if small compared to the batch this will follow a binomial distribution.\\
+    \\
+    An example of an acceptance sampling plan is to reject a sample (of size 50) if their are more than 2 non-conforming items. This plan would be denoted as:
+    $$n = 50, r = 3$$
+    or
+    $$n = 50, a = 2$$
+    where $r$ is reject when $>=$ the value, and $a$ is accept when $<=$ the value, and $n$ is the sample size.\\
+    \\
+    For any sampling play their is always the risk or rejecting a good batch (Type I error) or accepting a bad batch (Type II error).
+An ideal sampling play has the large probabilty of accepting a good batch, and a small probabilty of accepting a bad batch.
+
+    \begin{example}
+    {
+        An acceptance sampling scheme consists of a sample of $m$ components being taken from a batch.
+The atch is rejected if $r$ or more non-conforming components are found in the sample, otherwise it is accepted.
+    }
+
+    \begin{step}{If $n = 30$ and $r = 2$, find the probabilty of accepting a batch containing $1\%, 2\%, 4\%, 7\%, 10\%, 15\%, 20\%$ non-conforming components}
+    \begin{center}
+    \begin{tabular}{c|c|c}
+    Percentage & Distribution & Probabilty \\
+    \hline
+    1\%  & $X \sim B(30, 0.01)$ & 0.964 \\
+    2\%  & $X \sim B(30, 0.02)$ & 0.879 \\
+    4\%  & $X \sim B(30, 0.04)$ & 0.661 \\
+    7\%  & $X \sim B(30, 0.07)$ & 0.369 \\
+    10\% & $X \sim B(30, 0.10)$ & 0.184 \\
+    15\% & $X \sim B(30, 0.15)$ & 0.048 \\
+    20\% & $X \sim B(30, 0.20)$ & 0.010
+    \end{tabular}
+    \end{center}
+    \end{step}
+
+    \begin{step}{Draw the operating characteristic}
+
+    \begin{center}
+    \begin{tikzpicture}
+    \begin{axis}[
+        xlabel={\% non-conforming},
+        ylabel={P(accept)},
+        xmin=0, xmax=20,
+        ymin=0, ymax=1,
+    ]
+    \addplot[smooth,color=red,mark=*]coordinates {
+    (0, 1)(1,0.964)(2,0.879)(4,0.661)(7,0.369)(10,0.184)(15,0.048)(20,0.010)
+    };
+    \end{axis}
+    \end{tikzpicture}
+    \end{center}
+
+    \end{step}
+    \end{example}
+
+    \subsection{Accepting a plan}
+    The operating characteristic summerizes the implications of a plan however it does not decide what plan should be used.
+If the sample size and acceptable level of risk is specified it is possible to derive a plan.
+For example say a manafacture wants their to be at least a 95\% chance of accepting a batch with 1\% non-conforming itesm where $n = 50$.
+
+    \begin{center}
+    \begin{tabular}{c|c}
+    $r$ & $P(X < r-1)$ \\
+    \hline
+    1 & $0.605$ \\
+    2 & $0.911$ \\
+    3 & $0.986$ \\
+    4 & $0.998$ \\
+    \end{tabular}
+    \end{center}
+
+    \noindent
+    So when $r = 1$ the chance of accepting a batch with 1\% of non-conforming items is 60.5\% which is less than the required 95\%.
+Plans were $r > 2$ all fit the requirement, however we also want to minimize the chance of accepting a bad batch so the manafacture should use $r = 3$.
+
+    \begin{example}
+    {
+        An manafacturer will accept a risk of not more than 10\% of a batch of items containing 2\% non-conforming begin rejected. A decision is to be made by examining a sample of 50 items.
+    }
+
+    \begin{step}{Find an appropiate decision procedure}
+    $$X \sim B(50, 0.02)$$
+
+    \begin{center}
+    \begin{tabular}{c|c}
+    $r$ & $1 - P(X < r-1)$ \\
+    \hline
+    1 & $0.636$ \\
+    2 & $0.264$ \\
+    3 & $0.078$ \\
+    \end{tabular}
+    \end{center}
+
+    $$r = 3$$
+    \end{step}
+
+    \end{example}
+
+    \subsection{Double sampling plans}
+    With a single sample it is very likely to detect a very good/bad batch however for a more intermediate batch it makes a large sample is required for more accuracy.
+    $$n = 30; a = 1, r = 3$$
+    $$n = 30; a = 4, r = 5$$
+    This is an example of a double sampling plan, If the first sample has exactly 2 non-conforming items a second sample is taken else it is accepted/rejected.
+
+    \subsection{Acceptance sampling by variable}
+    Acceptance sampling can be carried out by measuring a varible, say the strength of each item.
+If the mean of theses measurements $\bar{x}$ is above a threshold $k$ then the batch is accepted, else it is rejected.
+The operating characteristic for a test by varible would be the mean of the measurement against the probabilty of accepting the batch.
+Its a reasonable to assume the normal distribution since we are conserned with the mean of a large sample.
+
+    \begin{example}
+    {
+        A batch of bricks are tested for strength which follows a normal distribution with a standard deviation of 5.5. A batch is accepted if the mean of five bricks is greated than 51.
+    }
+
+    \begin{step}{Draw the operating characteristic}
+    The standard deviation of the sample mean will be $\dfrac{5.5}{\sqrt{5}} = 2.46$. Since most data is within 3 standard deviation (on both sides) we will use a range of 44-58.
+
+    \begin{center}
+    \begin{tabular}{c|c|c}
+    $\mu$ & $\dfrac{51-\mu}{\dfrac{5.5}{\sqrt{5}}}$ & $P(\text{accept})$ \\
+    \hline
+    44 & 2.846 & 0.002 \\ 
+    46 & 2.033 & 0.021 \\
+    48 & 1.220 & 0.111 \\
+    50 & 0.407 & 0.342 \\
+    52 & -0.407 & 0.658 \\
+    54 & -1.220 & 0.889 \\
+    56 & -2.033 & 0.979 \\
+    58 & -2.846 & 0.998 \\
+    \end{tabular}
+    \end{center}
+
+    \begin{center}
+    \begin{tikzpicture}
+    \begin{axis}[
+        xlabel={$\mu$},
+        ylabel={P(accept)},
+        xmin=44, xmax=58,
+        ymin=0, ymax=1,
+    ]
+    \addplot[smooth,color=red,mark=*]coordinates {
+    (44, 0.002)(46, 0.021)(48, 0.111)(50, 0.342)(52, 0.658)(54, 0.889)(56, 0.979)(58, 0.998)
+    };
+    \end{axis}
+    \end{tikzpicture}
+    \end{center}
+
+    \end{step}
+
+    \begin{step}{The customer requires the probability of accepting a batch with a mean strength of 47 or less should be less than 10\%, would the customer be satisfied with the curren acceptance sampling}
+    47 is approximatly 0.05 or 5\% (from the operating characteristic) so the customer would be satisfied.
+    \end{step}
+
+    \begin{step}{Whats the smallest value for $k$ which would meet the customers requirements if $n = 6$}
+    \begin{align*}
+    \frac{k - 47}{\dfrac{5.5}{\sqrt{5}}} &> 1.2816 \\
+    k - 47 &> 2.8776 \\
+    k &> 49.8776
+    \end{align*}
+    $$k = 49.9$$
+    \end{step}
+
+    \end{example}

### 2017-05-31 src/mstats3.tex
Removed new pages and created files for MS03 and MS04.

\cchapter{Maths Statistics 3}

\newpage
\section{Geometric and Exponential Distributions}

### 2017-05-31 src/mstats3.tex
Added section titles.
Added geometic distributions section.

@@
 
-\newpage
 \section{Geometric and Exponential Distributions}
+
+    \subsection{Geometric Distributions}
+        Geometric distribution is like a binomial distribution however, instead of counting the number of sucesses, it counts the number of trials until a success is obtained.
+        \\\\
+        Conditions for a Geometric Distribution:
+        \begin{itemize}
+        \item Only two possible outcomes (success or failure).
+        \item Trials are independent.
+        \item Probabilty of success is the same for each trial.
+        \end{itemize}
+        The probability density function for a Geometric Distribution is as follows:
+        \begin{align*}
+        P(x) &= p(1-p)^{x-1}
+        \end{align*}
+        Which makes sence since the probability of a success is the $x$ times the probability of failure times the single probability of success.
+        \\\\
+        The mean of a Geometric Distribution is dirivied as follows.
+        \begin{align*}
+        \text{Let }q &= 1-p                     \\
+        E(X) &= \sum{x_i p_i}                   \\
+             &= p + 2pq + 3pq^2 + 4pq^3 + ...   \\
+             &= p(1 + 2q + 3q^2 + 4q^3 + ...)   \\
+             &= p(1-q)^{-2}                     \\
+             &= \dfrac{p}{p^2}                  \\
+             &= \dfrac{1}{p}
+        \end{align*}
+        \\\\
+        And the varience can be found using $E(X^2) - E(X)^2 = Var(X)$
+        \begin{align*}
+        E(X^2) &= \dfrac{2-p}{p^2} \text{  (using the method above)}    \\
+        Var(X) &= \dfrac{2-p}{p^2} - \left(\dfrac{1}{p}\right)^2        \\
+               &= \dfrac{2-p-1}{p^2}                                    \\
+               &= \dfrac{1-p}{p^2}                                      \\
+        \end{align*}
+    \subsection{Exponential Distributions}
+        (See Exponential Distribution section)
+
+\section{Estimators}
+\section{Estimation}
+\section{Hypothesis Testing}
+\section{Chi-Squared ($\chi^2$) Goodness of Fit Testing}

### 2017-06-01 src/mstats3.tex
Added estimators section.

@@
         \end{align*}
+
+        \begin{example}
+        {
+            An unbiased tetrahedral die has faces marked 1 to 4. When it is thrown on a table the score is the number on the face that is in contact with the table.
+        }
+
+            \begin{step}{Calculate the probability that it takes more than two throws to obtain a score of 4}
+                \begin{align*}
+                    p &= \dfrac{1}{4}                                                        \\
+                    q &= \dfrac{3}{4}                                                        \\
+                P(X > 2) &= 1 - P(X \leq 2)                                                     \\
+                        &= 1 - \left(\dfrac{1}{4} + \dfrac{1}{4} \times \dfrac{3}{4}\right)    \\
+                        &= \dfrac{9}{16}
+                \end{align*}
+            \end{step}
+
+            \begin{step}{The number of throws, $Y$, that it takes for two diffrent scores to occor at least once is given by $Y = 1 + X$, where $X$ has a geometric distribution with parameter $\dfrac{3}{4}$.
+Determine values for $E(Y)$ and $Var(Y)$.}
+                \begin{align*}
+                E(Y) &= E(1+X)                          \\
+                    &= E(1) + E(X)                     \\
+                    &= 1 + \dfrac{1}{\frac{4}{3}}      \\
+                    &= \dfrac{7}{3}                    
+                \end{align*}
+
+                \begin{align*}
+                Var(Y) &= Var(1+X)                                  \\
+                    &= Var(1) + Var(X)                           \\
+                    &= 0 + \dfrac{\frac{1}{4}}{(\frac{3}{4})^2}  \\
+                    &= \dfrac{4}{9}
+                \end{align*}
+            \end{step}
+        \end{example}
+
     \subsection{Exponential Distributions}
@@
 \section{Estimators}
+    \paragraph{Sample Statistic} A function of observed data, used to estimate parameters (such as mean and varience).
+
+    \paragraph{Sampling Distribution} The probability distribution for a statistic based on random samples.
+
+    \paragraph{Population Parameter} A statistical measure for a population, used within a general distribution function such a the normal distribution.
+
+    \paragraph{Estimators} A function to estimate a population parameter from some samples.
+
+    \paragraph{Estimates} The value yielded by an estimator.
+
+    \paragraph{Unbiased estimator} If $T$ is the estimator for a parameter $\theta$ and $E(T) = \theta$ then $T$ is an unbiased estimator for $\theta$.
+
+    \paragraph{Consistency} Estimators are consistent if their variation approaches zero as the sample size approaches infinity.
+
+    \paragraph{Efficiency} A measure of quaility of an estimator, a more effiecent estimator needs fewer observations than a less efficent one.
+
+    \paragraph{Relative Efficiency} The ratio of two procedures effciencies, given by $RE(T_1 : T_2) = \dfrac{Var(T_2)}{Var(T_1)}$
+
+    \begin{example}
+    {
+        The conductivity, $\gamma$, of metal wire is estimated by observing a related random variable, R, which has probability density function
+
+        $$
+        f(r) = 
+        \begin{cases}
+        \dfrac{2r}{\gamma^2} & 0 \leq r \leq \gamma \\
+        0 & \text{otherwise}
+        \end{cases}
+        $$
+    }
+
+        \begin{step}{Show that $\frac{3}{2}R$ is an unbiased estimator of $\gamma$}
+            \begin{align*}
+            E\left(\dfrac{3}{2}R\right) &= \dfrac{3}{2} \times E(R)                                             \\
+                                        &= \dfrac{3}{2} \times \int_0^\gamma r \times \dfrac{2r}{\gamma^2} dr   \\
+                                        &= \int_0^\gamma \dfrac{3r^2}{\gamma^2} dr                              \\
+                                        &= \left[\dfrac{r^3}{\gamma^2}\right]^\gamma_0                          \\
+                                        &= \dfrac{\gamma^3}{\gamma^2} - \dfrac{0}{\gamma^2}                     \\
+                                        &= \gamma 
+            \end{align*}
+        \end{step}
+
+        \begin{step}{Given that the variance of $R$ is $\frac{1}{18}\gamma^2$, find, in terms of $\gamma$, the variance of $\frac{3}{2}R$}
+            \begin{align*}
+            Var\left(\frac{3}{2}R\right) &= \left(\dfrac{3}{2}\right)^2 \times Var(R)   \\
+                                        &= \dfrac{9}{4} \times \dfrac{\gamma^2}{18}      \\
+                                        &= \dfrac{\gamma^2}{8}
+            \end{align*}
+        \end{step}
+
+        \begin{step}{The conductivity can also be estimated by making observations of a random varible, S, which has mean $\frac{1}{4}\gamma$ and $\frac{1}{16}\gamma^2$.
+The ranomd varible $T$ is defined by $T = S_1 + S_2 + S_3$, where $S_1$, $S_2$ and $S_3$ are three independent observations of $S$.\\\\Find the value of the constant $k$ such that $kT$ is an unbiased estimator of $\gamma$}
+            \begin{align*}
+                \gamma &= E(kT)                                                                 \\
+                \gamma &= k \times E(T)                                                         \\
+                \gamma &= k \times E(S_1 + S_2 + S_3)                                           \\
+                \gamma &= k \times \frac{1}{4}\gamma + \frac{1}{4}\gamma + \frac{1}{4}\gamma    \\
+                \gamma &= k \times \frac{3}{4}\gamma                                            \\
+                k &= \frac{4}{3}
+            \end{align*}
+        \end{step}
+
+        \begin{step}{Find the relative efficiency of $\frac{3}{2}R$ with respect to $kT$}
+            \begin{align*}
+            Var\left(\frac{4}{3}T\right) &= \left(\dfrac{4}{3}\right)^2 \times Var(T)       \\
+                                         &= \dfrac{16}{9} \times Var(S_1 + S_2 + S_3)       \\
+                                         &= \dfrac{16}{9} \times (\dfrac{1}{16}\gamma^2 + \dfrac{1}{16}\gamma^2 + \dfrac{1}{16}\gamma^2)                             \\
+                                         &= \dfrac{16}{9} \times \dfrac{3}{16}\gamma^2      \\
+                                         &= \dfrac{\gamma^2}{3}
+            \end{align*}
+            \begin{align*}
+                RE\left(\frac{3}{2}R : \frac{4}{3}T\right) &= \dfrac{Var\left(\dfrac{4}{3}T\right)}{ Var\left(\dfrac{3}{2}R\right)} \\
+                &= \dfrac{\dfrac{\gamma^2}{3}}{\dfrac{\gamma^2}{8}}      \\
+                &= \dfrac{\gamma^2}{3} \times \dfrac{8}{\gamma^2}        \\
+                &= \dfrac{8}{3}
+            \end{align*}
+        \end{step}
+
+        \begin{step}{State, with justification, which of $\frac{3}{2}R$ and $kT$ is a preferred estimator of $\gamma$}
+            $\dfrac{8}{3} > 1$ so $\dfrac{3}{2}R$ is the preffered estimator.
+        \end{step}
+
+    \end{example}
+
+    \begin{example}
+    {
+        The random variable $X$ denotes the number of successes in a random sample of size $n$ from a large population in which the proportion of successes is $p$.
+        \\\\
+        The random varible $Y$ denotes the number of successes in an independent random sample of size $3n$ from the same large population.
+    }
+
+    \begin{step}{Write down expressions for the mean and the variance of each of $X$ and $Y$}
+        X: Mean = $np$, Variance = $np(1-p)$
+
+        Y: Mean = $3np$, Variance = $3np(1-p)$
+    \end{step}
+
+    \begin{step}
+    {
+        Two estimators suggested for $p$ are:
+        \begin{center}
+        $\hat{P}_1 = \dfrac{X + Y}{4n}$ and $\hat{P}_2 = \dfrac{1}{2}\left(\dfrac{X}{n} + \dfrac{Y}{3n}\right)$
+        \end{center}
+        Show that both $\hat{P}_1$ and $\hat{P}_2$ are unbiased and consistent estimators for $p$, and that $\hat{P}_1$ is the more efficent.
+    }
+
+    \begin{align*}
+        E\left(\dfrac{X + Y}{4n}\right) &= \dfrac{np + 3np}{4n}     \\
+        &= \dfrac{4np}{4n}                                          \\
+        &= p
+    \end{align*}
+
+    \begin{align*}
+        E\left(\dfrac{1}{2}\left(\dfrac{X}{n} + \dfrac{Y}{3n}\right)\right) &= \frac{1}{2} \times \left(\dfrac{np}{n}+\dfrac{3np}{3n}\right)    \\
+        &= \frac{1}{2} \times (p + p) \\
+        &= p
+    \end{align*}
+
+    \begin{align*}
+        Var\left(\dfrac{X + Y}{4n}\right) &= \dfrac{npq + 3npq}{16n^2}      \\
+        &= \dfrac{4pq}{16n}                                                 \\
+        &= \dfrac{pq}{4n}                                                   \\
+        &= \dfrac{p(1-p)}{4n}
+    \end{align*}
+
+    \begin{center}
+    as $Var\left(\dfrac{X + Y}{4n}\right) \rightarrow 0, n \rightarrow \infty$
+    \end{center}
+
+    \begin{align*}
+        Var\left(\dfrac{1}{2}\left(\dfrac{X}{n} + \dfrac{Y}{3n}\right)\right) &= \frac{1}{2^2} \times \left(\dfrac{npq}{n^2}+\dfrac{3npq}{9n^2}\right)        \\
+        &= \frac{1}{4} \times \dfrac{12npq}{9n^2}   \\
+        &= \dfrac{pq}{3n}                           \\
+        &= \dfrac{p(1-p)}{3n}
+    \end{align*}
+
+    \begin{center}
+    as $Var\left(\dfrac{1}{2}\left(\dfrac{X}{n} + \dfrac{Y}{3n}\right)\right) \rightarrow 0, n \rightarrow \infty$
+    \end{center}
+
+    \begin{align*}
+    RE(\hat{P}_1 : \hat{P}_2) &= \dfrac{Var(\hat{P}_2)}{Var(\hat{P}_1)}  \\
+    &= \dfrac{\dfrac{p(1-p)}{3n}}{\dfrac{p(1-p)}{4n}}                               \\
+    &= \frac{4}{3}                                                                  \\
+    \end{align*}
+
+    $\dfrac{4}{3} > 1$ so $\hat{P}_1$ is the best estimate for parameter $p$.
+
+    \end{step}
+
+    \end{example}
+
 \section{Estimation}

### 2017-06-01 src/mstats3.tex
Changed wording in example.
Added proof of E(S^2) = sigma^2.
Added estimation and hypothesis section.

@@
 \section{Estimators}
+    \subsection{Estimators, Unbiased, consistent and efficiency}
         \paragraph{Sample Statistic} A function of observed data, used to estimate parameters (such as mean and varience).
@@
 
-    $\dfrac{4}{3} > 1$ so $\hat{P}_1$ is the best estimate for parameter $p$.
+        $\dfrac{4}{3} > 1$ so $\hat{P}_1$ is the more efficent.
 
@@
 
+    \subsection{Proof of $E(S^2) = \sigma^2$}
+        Using these methods, we can now prove $E(S^2) = \sigma^2$
+        \begin{align*}
+            \bar{X}  &= \sum{\dfrac{X_i}{n}} \\
+            n\bar{X} &= \sum{X_i}\\\\
+            \dfrac{\sigma^2}{n} &= E(\bar{X}^2) - \mu^2 \\
+            E(\bar{X}^2) &= \dfrac{\sigma^2}{n} + \mu^2
+        \\\\
+        E(S^2) &= E\left(\dfrac{1}{n-1} \times \sum{(X_i - \bar{X})^2}\right)               \\
+            &= \dfrac{1}{n-1} \times E\left(\sum{(X_i - \bar{X})^2}\right)               \\
+            &= \dfrac{1}{n-1} \times \sum{E((X_i - \bar{X})^2)}                          \\
+            &= \dfrac{1}{n-1} \times \sum{E(X_i^2 - 2\bar{X}X_i + \bar{X}^2)}            \\
+            &= \dfrac{1}{n-1} \times \sum{(E(X_i^2) - E(2\bar{X}X_i) + E(\bar{X}^2))}    \\
+            &= \dfrac{1}{n-1} \times \left(\sum{E(X_i^2)} - \sum{E(2\bar{X}X_i)} + \sum{E(\bar{X}^2))}\right) \\
+            &= \dfrac{1}{n-1} \times \left(n(\sigma^2 - \mu^2) - E\left(\sum{2\bar{X}X_i}\right) + E\left(\sum{\bar{X}^2}\right)\right)\\
+            &= \dfrac{1}{n-1} \times \left(n(\sigma^2 - \mu^2) - E\left(2\bar{X} \times n\bar{X}\right) + E\left(n\bar{X}^2\right)\right)\\
+            &= \dfrac{1}{n-1} \times \left(n(\sigma^2 - \mu^2) - E\left(2n\bar{X}^2\right) + E\left(n\bar{X}^2\right)\right)\\
+            &= \dfrac{1}{n-1} \times \left(n(\sigma^2 - \mu^2) - E\left(n\bar{X}^2\right)\right)\\
+            &= \dfrac{1}{n-1} \times \left(n(\sigma^2 - \mu^2) - nE\left(\bar{X}^2\right)\right)\\
+            &= \dfrac{1}{n-1} \times \left(n(\sigma^2 - \mu^2) - n\left(\dfrac{\sigma^2}{n} + \mu^2\right)\right)\\
+            &= \dfrac{1}{n-1} \times (n\sigma^2 - n\mu^2 - \sigma^2 + n\mu^2)\\
+            &= \dfrac{1}{n-1} \times (n\sigma^2 - \sigma^2)\\
+            &= \dfrac{1}{n-1} \times \sigma^2(n-1)\\
+            &= \sigma^2
+        \end{align*}
+
 \section{Estimation}
+    \subsection{Confidence interval for diffrence in means of two normal distributions}
+        The confidence interval for the diffrence of two means is as simple as finding the diffrence between the two sets of samples and finding the confidence interval as normal.
+
+        \begin{example}
+        {
+            The following is the blood pressure of patients while sitting and standing.
+
+            \begin{center}
+            \begin{tabular}{l|c|c|c|c|c|c|c|c|c|c|c|c}
+                Patient     & A & B & C & D & E & F & G & H & I & J & K & L \\
+                \hline
+                Sitting     & 72 & 79 & 69 & 76 & 82 & 66 & 87 & 78 & 85 & 77 & 84 & 70 \\
+                Standing    & 81 & 70 & 79 & 84 & 88 & 77 & 79 & 85 & 91 & 74 & 88 & 80 
+            \end{tabular}
+            \end{center}
+
+            Construct a 95\% confidence inteval for the mean diffrence between the patient's blood pressure assuming the diffrence in blood pressure are normally distributed, and state weather the claim that patients blood pressure are diffrent is supported.
+        }
+
+        \begin{step}{Find the diffrence}
+            $$
+                d_i = 9, -9, 10, 8, 6, 11, -8, 7, 6, -3, 4, 10
+            $$
+        \end{step}
+
+        \begin{step}{Find the mean and variance of the diffrences}
+            $$
+                \bar{d} = 4.5
+            $$
+            $$
+                s_d = 7.01
+            $$
+        \end{step}
+
+        \begin{step}{Find the critical value}
+            $$
+                cv = t_{11} = 2.201
+            $$
+        \end{step}
+
+        \begin{step}{Construct the confindence inteval}
+            \begin{align*}
+                4.25 \pm 2.201 \times \dfrac{7.01}{\sqrt{12}} &= (-0.20, 8.70)
+            \end{align*}
+        \end{step}
+
+        \begin{step}{Conclude}
+            The confindence interval includes 0, so their is no evidence to support the claim that patients blood pressure is diffrent between sitting and standing.
+        \end{step}
+
+        \end{example}
+
+    \subsection{Confindence interval for variance}
+        (See SS05 Confindence interval for variance)
+
+    \subsection{Confindence interval for the ratio of two variances}
+         The ratio of variances is used it ANOVA and variance test's, but we can also find its confindence interval.
+
+         Given $s_x$, $n_x$, $s_y$ and $n_y$ of two normal distributions, the confidence interval for the ratio of the variances is as follows:
+         $$
+         \dfrac{1}{F_{v_1, v_2}} \leq \frac{VR}{F_{\text{calc}}} \leq F_{v_2, v_1}
+         $$
+         Where $F_{\text{calc}} = \dfrac{s_x^2}{s_y^2}$, $v_1 = n_x - 1$ and $v_2 = n_y - 1$
+
+         \begin{example}
+         {
+             The broadband speed of 12 rural villages are as follows:
+             $$1.9, 2.6, 1.8, 3.4, 2.2, 3.0, 2.7, 3.7, 2.7, 1.9, 3.4, 3.1$$
+             The broadband speed of 9 small towns are as follows:
+             $$7.8, 7.7, 7.5, 7.7, 8.0, 7.3, 7.7, 7.4, 7.8$$
+
+             Determine a 99\% confindence interval for the variance ratio $\dfrac{\sigma_X^2}{\sigma_Y^2}$, and comment on the suggestion that rutal villages is more variable than small towns.
+         }
+
+         \begin{step}{Find the variance of $X$ and $Y$}
+            $$s_X^2 = 0.41636$$
+            $$s_Y^2 = 0.04778$$
+         \end{step}
+
+         \begin{step}{Find the critical values}
+            \begin{align*}
+            v_1       &= 12 - 1   \\
+                      &= 11       \\
+            v_2       &= 9 - 1    \\
+                      &= 8        \\
+            F_{11, 8} &= 7.104    \\
+            F_{8, 11} &= 5.682
+            \end{align*}
+            $$\dfrac{1}{7.104} \leq \dfrac{VR}{8.7146} \leq 7.682$$
+            $$1.23 \leq VR \leq 49.5$$
+         \end{step}
+
+         \begin{step}{Conclude}
+            1 is not inside the confidence interval, so broadband speeds are more varible in villages.
+         \end{step}
+
+         \end{example}
+
 \section{Hypothesis Testing}
+    \subsection{Test for diffrence between in mean}
+        (See SS5 Test for diffrence between in mean)
+
+    \subsection{Tests for a variance}
+        (See SS5 Test for variance equality)
+
+    \subsection{Test for the ratio of two variances}
+        Much like how we can test for variace equality, we can test the ratio of two variances in a simular mannor.
+
+        \subsubsection{Process}
+            \begin{enumerate}
+            \item Hypothesis.
+            \item Find the sample variances
+            \item Calculate the $F_{\text{test}}$ divided by the ratio that we assume in $H_0$
+            \item Determine the critical value
+            \item Conclude
+            \end{enumerate}
+
+        \begin{example}
+        {
+            With the following data:
+            \begin{center}
+            \begin{tabular}{c|c|c}
+            Distribution & $s^2$ & $n$ \\
+            $S$ & 3372.75 & 12 \\
+            $N$ & 290.5 & 8 \\
+            \end{tabular}
+            \end{center}
+            Test the claim that the standard deviation of $S$ is twice that of $N$.
+        }
+
+        \begin{step}{Hypothesis}
+            $$ H_0: \sigma_S = 2 \sigma_N $$
+            $$ H_0: \sigma_S \ne 2 \sigma_N $$
+        \end{step}
+
+        \begin{step}{Sample variances}
+            $$s_S^2 = 3372.75$$
+            $$s_N^2 = 290.5$$
+        \end{step}
+
+        \begin{step}{Calculate $F_\text{test}$}
+            \begin{align*}
+                F_\text{test} &= \dfrac{3372.75}{2^2 \times 290.5}  \\
+                              &= 2.90
+            \end{align*}
+        \end{step}
+
+        \begin{step}{Critical value}
+            \begin{align*}
+            v_1 &= 12 - 1       \\
+                &= 11           \\
+            v_2 &= 8 - 1        \\
+                &= 7            \\\\
+            F_{11, 7} &= 3.60
+            \end{align*}
+        \end{step}
+
+        \begin{step}{Conclude}
+            $2.80 < 3.60$, so accept $H_0$ significant evidence that the standard deviation of $S$ is twice that of $N$.
+        \end{step}
+
+        \end{example}
+
 \section{Chi-Squared ($\chi^2$) Goodness of Fit Testing}

### 2017-06-03 src/mstats3.tex
Added goodness of fit.
Completed all but 1 section of MS03.

--- 2017-05-31 src/mstats3.tex
@@
 
-\section{Geometric and Exponential Distributions}
+    \section{Further Probability}
+        \subsection{Bayes' Theorem}
+            Bayes' theorem states:
+            $$
+            P(A_j|B) = \dfrac{P(A_j)P(B|A_j)}{\displaystyle\sum^n_{i=1}{P(A_i)P(B|A_i)}}
+            $$
 
-    \subsection{Geometric Distributions}
-        Geometric distribution is like a binomial distribution however, instead of counting the number of sucesses, it counts the number of trials until a success is obtained.
+            \begin{example}
+            {
+                Three members of a golf club are running for captain.
+
+                \begin{center}
+                    \begin{tabular}{c|l|c|c}
+                        Event & Person  & Probability of election & Probability of increasing fees \\
+                        \hline
+                        $A_1$ & Alan    & 0.3 & 0.8 \\
+                        $A_2$ & Brian   & 0.5 & 0.1 \\
+                        $A_3$ & Clive   & 0.2 & 0.4 \\
+                    \end{tabular} 
+                \end{center}
+
+                After the election, fees were increased. Find the probability Clive was elected. 
+            }
+
+            \begin{step}{Use Bayes' Theorem}
+                Let $B$ be the probability that their is a fee increase.
+                \begin{align*}
+                    P(A_3|B) &= \dfrac{P(A_3)P(B|A_3)}{P(A_1)P(B|A_1)+P(A_2)P(B|A_2)+P(A_3)P(B|A_3)}    \\
+                             &= \dfrac{0.2\times0.4}{0.3\times0.8+0.5\times0.1+0.2\times0.4}            \\
+                             &= \frac{8}{37}
+                \end{align*}
+            \end{step}
+
+            \end{example}
+
+    \section{Linear Combinations of Random Variables}
+        \subsection{Mean and variance of linear combination of random variables}
+        Covariance tells you how two random varibles vary togther. Its function is defined as the following:
+        $$
+        Cov(X, Y) = E((X-\mu_x)(Y-\mu_Y)) = E(XY) - \mu_x \mu_y
+        $$
+        Using the covariance we can find the product moment correlation coefficent of two random varibles with the following:
+        $$
+        \rho = \dfrac{Cov(X, Y)}{\sigma_X \sigma_Y}
+        $$
+        Additionaly the variance can be computed with the following:
+        $$
+        Var(aX \pm bY) = a^2 Var(X) + b^2 Var(Y) \pm 2ab Cov(X, Y)
+        $$
+
+        \begin{example}
+        {
+            \begin{center}
+                \begin{tabular}{c|c|c|c|c|c|c}
+                    X/Y & 2 & 3 & 4 & 5 & 6 & $P(Y = y)$\\
+                    \hline
+                    3 & 0.00 & 0.05 & 0.20 & 0.20 & 0.05 & 0.50 \\
+                    4 & 0.00 & 0.15 & 0.10 & 0.05 & 0.00 & 0.30 \\
+                    5 & 0.05 & 0.05 & 0.10 & 0.00 & 0.00 & 0.20 \\
+                    \hline
+                    $P(X = x)$ & 0.05 & 0.25 & 0.40 & 0.25 & 0.05 & 1.00 \\
+                \end{tabular}
+            \end{center}
+        }
+
+            \begin{step}{State why $E(X)=4$ and show that $Var(X)=0.9$}
+                The distribution $X$ is symetrical around 4, thus $E(X) = 4$
+                \begin{align*}
+                    E(X^2) &= 2^2 \times 0.05 + 3^2 \times 0.25 + 4^2 \times 0.40 + 5^2 \times 0.25 + 6^2 \times 0.05\\
+                    &= 16.9 \\
+                    Var(X) &= 16.9 - 4^2 \\
+                    &= 0.9
+                \end{align*}
+            \end{step}
+
+            \begin{step}{Given that $$E(Y)=3.7,Var(Y)=0.61,E(XY)=14.4$$Calculate values for $Cov(X,Y)$ and $\rho_{XY}$}
+                \begin{align*}
+                Cov(X,Y)  &= 14.4 - 4 \times 3.7                            \\
+                        &= -0.4                                           \\
+                \rho_{XY} &= \dfrac{-0.4}{\sqrt{0.9} \times \sqrt{0.61}}    \\
+                        &= -0.540
+                \end{align*}
+            \end{step}
+
+            \begin{step}{Find the mean and variance of $T = X + Y$}
+                \begin{align*}
+                    E(T)    &= E(X) + E(Y)                  \\
+                            &= 4 + 3.7                      \\
+                            &= 7.7                          \\
+                    Var(T)  &= 0.9 + 0.61 + 2 \times -0.4   \\
+                            &= 0.71
+                \end{align*}
+            \end{step}
+
+            \begin{step}{Find the mean and variance of $D = X - Y$}
+                \begin{align*}
+                    E(D)    &= E(X) - E(Y)                  \\
+                            &= 4 - 3.7                      \\
+                            &= 0.3                          \\
+                    Var(D)  &= 0.9 + 0.61 - 2 \times -0.4   \\
+                            &= 2.31
+                \end{align*}
+            \end{step}
+
+        \end{example}
+
+        \subsection{Mean and variance of linear combination of independent random variables}
+            (See SS04 Combination of random varibles)
+
+        \subsection{Linear combination of independent normal variables}
+            (See SS04 Combination of random varibles)
+
+    \section{Distributional Approximations}
+        \subsection{Mean and variance of binomial and poisson distributions}
+            \subsubsection{Proof of $\mu = np$ for the Binomial distribution}
+                \begin{align*}
+                    \mu &= E(X)                                                                     \\
+                        &= \sum x_i p_i                                                     \\
+                        &= \sum^n_{x=0} x \times \dfrac{n!}{x!(n-x)!} \times p^x(1-p)^{n-x}       \\
+                        &= \sum^n_{x=1} x \times \dfrac{n!}{x!(n-x)!} \times p^x(1-p)^{n-x} && \text{Summation is zero at $x=0$, so lower bound becomes $x=1$} \\
+                        &= np \sum^n_{x=1} \dfrac{(n-1)!}{(x-1)!(n-x)!} \times p^{x-1}(1-p)^{n-x} && \text{Take of factor of $np$ and cancel $x$ with $x!$} \\
+                        &= np \sum^{n-1}_{y=0} \dfrac{(n-1)!}{y!(n-y-1)!} \times p^y(1-p)^{n-y-1} && \text{Substitute $y=1$} \\
+                        &= np \sum^{n-1}_{x=1} B(n-1, p) && \text{The expression is the equivilent of $B(n-1, p)$} \\
+                        &= np && \text{Sum of all binomial terms = 1}
+                \end{align*}
+
+            \subsubsection{Proof of $\sigma^2 = n(n-1)p$ for the Binomial distribution}
+                \begin{align*}
+                    E(X(X-1)) &= E(X^2) - E(X)                  && \text{Expand}\\
+                              &= E(X^2) - np                    \\
+                              &= E(X^2) - E(X)^2 + E(E)^2 - np  && \text{The trick is to add and subtract $E(X)^2$}\\
+                              &= \sigma^2 + n^2p^2 - np     \\            
+                     \sigma^2 &= E(X(X-1)) - n^2p^2 + np
+                    \\\\\\
+                    E(X(X-1)) &= \sum^n_{x=0} x \times (x-1) \times \dfrac{n!}{x!(n-x)!} \times p^x(1-p)^{n-x} \\
+                              &= \sum^n_{x=2} x \times (x-1) \times \dfrac{n!}{x!(n-x)!} \times p^x(1-p)^{n-x} && \text{Value at 0 and 1 is 0 so lower limit becomes $x=2$} \\
+                              &= n(n-1)p^2 \sum^n_{x=2} \dfrac{(n-2)!}{(x-2)!(n-x)!} \times p^{x-2}(1-p)^{n-x} && \text{Take out a factor of $n(n-1)p^2$ and cancel $x(x-1)$} \\
+                              &= n(n-1)p^2 \sum^{n-2}_{y=0} \dfrac{(n-2)!}{y!(n-2-y)!} \times p^y(1-p)^{n-2-y} && \text{Substitute $y=x-2$} \\
+                              &= n(n-1)p^2 \sum^{n-2}_{x=2} B(n-2, p) && \text{The expression is the equivilent of $B(n-2, p)$} \\
+                              &= n(n-1)p^2 && \text{Sum of all binomial terms = 1}
+                    \\\\\\
+                    \sigma^2 &= E(X(X-1)) - n^2p^2 + np \\
+                             &= n(n-1)p^2 - n^2p^2 + np && \text{Substiute in the value we derived above} \\
+                             &= n^2p^2 - np^2 - n^2p^2 + np \\
+                             &= np - np^2   \\
+                             &= np(1-p)
+                \end{align*}
+
+            \subsubsection{Proof of $\mu=\lambda$ for the Poisson distribution}
+                \begin{align*}
+                    \mu &= \sum x_i p_i \\
+                        &= \sum_{x=0}^\infty x \dfrac{e^{-\lambda}\times\lambda^{x}}{x!} \\
+                        &= \sum_{x=1}^\infty x \dfrac{e^{-\lambda}\times\lambda^{x}}{x!} && \text{Value at 0 is 0 so lower limit becomes $x=1$} \\
+                        &= \lambda \sum_{x=1}^\infty \dfrac{e^{-\lambda}\times\lambda^{x-1}}{(x-1)!} && \text{Factor out $\lambda$ and cancel out $x$} \\
+                        &= \lambda \sum_{y=0}^\infty \dfrac{e^{-\lambda}\times\lambda^{y}}{y!} && \text{Substitute $y = x - 1$} \\
+                        &= \lambda && \text{Sum of all poisson terms = 1}
+                \end{align*}
+
+            \subsubsection{Proof of $\sigma^2=\lambda$}
+                \begin{align*}
+                    E(X(X-1)) &= E(X^2) - E(X)                  && \text{Expand}\\
+                              &= E(X^2) - \lambda                    \\
+                              &= E(X^2) - E(X)^2 + E(E)^2 - \lambda  && \text{The trick is to add and subtract $E(X)^2$}\\
+                              &= \sigma^2 + \lambda^2 - \lambda     \\            
+                     \sigma^2 &= E(X(X-1)) - \lambda^2 + \lambda
+                    \\\\\\
+                    E(X(X-1)) &= \sum^\infty_{x=0} x(x-1) \dfrac{e^-\lambda \times \lambda^x}{x!} \\
+                              &= \sum^\infty_{x=2} x(x-1) \dfrac{e^-\lambda \times \lambda^x}{x!} && \text{Value at 0 and 1 is zero so lower limit becomes $x=2$}\\
+                              &= \lambda^2 \sum^\infty_{x=2} \dfrac{e^-\lambda \times \lambda^{(x-2)}}{(x-2)!} && \text{Factor out $\lambda^2$ and cancel $x(x-1)$} \\
+                              &= \lambda^2 \sum^\infty_{y=0} \dfrac{e^-\lambda \times \lambda^y}{y!} && \text{Substitute $y=x-2$} \\
+                              &= \lambda^2 && \text{Sum of poisson terms = 1}
+                    \\\\\\
+                    \sigma^2 &= \lambda^2 - \lambda^2 + \lambda \\
+                             &= \lambda
+                \end{align*}
+
+        \subsection{Approximating binomial using the Possion distribution}
+            (See SS04 Approximating binomial using the Possion distribution)
+
+        \subsection{Approximating binomial using the normal distribution}
+            (See SS04 Approximating binomial using the normal distribution)
+
+        \subsection{Approximating Possion using the normal distribution}
+            (See SS04 Approximating Possion using the normal distribution)
+
+    \section{Estimation}
+        \subsection{Estimation of sample sizes to achieve required width confidence interval}
+            (Couldnt find any questions, probably simple though)
+        \subsection{Confindence interval for the diffrence between means of two normal random variables}
+            As simple as:
+            $$
+            \bar{x} - \bar{y} \pm z \sqrt{\dfrac{s^2_x}{n_x} + \dfrac{s^2_y}{n_y}}
+            $$
+            (May require approximation)
+
+            \begin{example}
+            {
+                A sample of 50 male Eastern Grey kangaroos had a mean weight of 42.6 kg and a standard deviation of 6.2 kg.
                 \\\\
-        Conditions for a Geometric Distribution:
-        \begin{itemize}
-        \item Only two possible outcomes (success or failure).
-        \item Trials are independent.
-        \item Probabilty of success is the same for each trial.
-        \end{itemize}
-        The probability density function for a Geometric Distribution is as follows:
+                A sample of 50 male Western Grey kangaroos had a mean weight of 39.7 kg and a standard deviation of 5.3 kg.
+                \\\\
+                Construct a 98\% confidence interval for the difference between the mean weight of male Eastern Grey kangaroos and that of male Western Grey kangaroos.
+            }
+                \begin{step}{Construct the confidence interval}
                     \begin{align*}
-        P(x) &= p(1-p)^{x-1}
+                        42.6 - 39.7 \pm 2.3263 \times \sqrt{\dfrac{6.2^2}{50} + \dfrac{5.3^2}{50}} &= (0.217, 5.58)
                     \end{align*}
-        Which makes sence since the probability of a success is the $x$ times the probability of failure times the single probability of success.
-        \\\\
-        The mean of a Geometric Distribution is dirivied as follows.
+                \end{step}
+            \end{example}
+
+        \subsection{Mean and variance of sample proportion}
+            The mean and variance of a sample proportion is as follows:
+            $$\mu = p$$
+            $$\sigma^2 = \dfrac{p(1-p)}{n}$$
+
+        \subsection{Unbiased estimator of population proportion}
+            $$\hat{p} = \frac{x}{n}$$
+
+        \subsection{Approximating sample proportion using the normal approximation}
+            \begin{example}
+            {
+                Of 480 men polled, 264 voted Yes and of 500 women polled, 220 voted Yes.
+            }
+
+                \begin{step}{Construct a 95\% confidence interval for the diffrence between the proportion of men and women who voted Yes}
                     \begin{align*}
-        \text{Let }q &= 1-p                     \\
-        E(X) &= \sum{x_i p_i}                   \\
-             &= p + 2pq + 3pq^2 + 4pq^3 + ...   \\
-             &= p(1 + 2q + 3q^2 + 4q^3 + ...)   \\
-             &= p(1-q)^{-2}                     \\
-             &= \dfrac{p}{p^2}                  \\
-             &= \dfrac{1}{p}
+                        \hat{p}_m &= \frac{264}{480} = 0.55                                     \\
+                        \hat{p}_w &= \frac{220}{500} = 0.44                                     \\
+                        M &\sim N(0.55, \dfrac{0.55 \times 0.45}{480}) \sim N(0.55, 0.000516)   \\
+                        W &\sim N(0.44, \dfrac{0.44 \times 0.55}{500}) \sim N(0.44, 0.000493)
                     \end{align*}
-        \\\\
-        And the varience can be found using $E(X^2) - E(X)^2 = Var(X)$
+
+                    $$(0.55 - 0.44) \pm 1.96 \times \sqrt{0.000516 + 0.000493} = (0.048, 0.172)$$
+                \end{step}
+
+                \begin{step}{Comment on the claim that men voted yes exceeded the percentage of women who voted yes by 2.5\%}
+                    The confidence interval is greater than 2.5\%, so their is evidence to support the claim.
+                \end{step}
+            \end{example}
+
+        \subsection{Confindence interval for population proportion}
+            As you would exspect:
+            $$
+            \hat{p} \pm z \times \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}
+            $$
+
+            \begin{example}
+            {
+                23 out of a sample of 200 pens fail to write.
+            }
+
+            \begin{step}{Calculate an approximate 96\% confidence interval of pens that fail to write.}
+                $$\hat{p} = \frac{23}{200} = 0.115$$
                 \begin{align*}
-        E(X^2) &= \dfrac{2-p}{p^2} \text{  (using the method above)}    \\
-        Var(X) &= \dfrac{2-p}{p^2} - \left(\dfrac{1}{p}\right)^2        \\
-               &= \dfrac{2-p-1}{p^2}                                    \\
-               &= \dfrac{1-p}{p^2}                                      \\
+                    0.115 \pm 2.0537 \times \sqrt{\dfrac{0.115 \times (1-0.115)}{200}} = (0.069, 0.161)
                 \end{align*}
-    \subsection{Exponential Distributions}
-        (See Exponential Distribution section)
+            \end{step}
 
-\section{Estimators}
-\section{Estimation}
+            \begin{step}{Comment on the claim that at most 2 in 50 pens fail.}
+                $\dfrac{2}{50} = 0.04$, 0.04 is less than the confidence interval, so their is evidence to reject their claim
+            \end{step}
+
+            \end{example}
+
+        \subsection{Confindence interval for mean of poisson distributions}
+            (See SS04 Confindence intervals for Possion)
+
+        \subsection{Confindence interval for the diffrence in poisson distribution means}
+            $$(\lambda_x - \lambda_y) \pm z \times \sqrt{\lambda_x + \lambda_y}$$
+
+            \begin{example}
+            {
+                In a sample of 26 weeks, 507 orders were received on Monday, and 416 orders were received on Friday.
+            }
+
+            \begin{step}{Construct a 99\% approximate confidence interval for $\lambda_M - \lambda_F$ where $\lambda_M$ and $\lambda_F$ represent the daily number of orders}
+                $$(507-416) \pm 2.5758 \times \sqrt{507+416} = (12.7, 169.3)$$
+                $$\dfrac{(12.7, 169.3)}{26} = (0.49, 6.51)$$
+
+            \end{step}
+
+            \begin{step}{Comment on the belief that more orders are received on mondays than fridays}
+                Confindence interval is above 0, therefore the belief is justified.
+            \end{step}
+
+            \end{example}
+
     \section{Hypothesis Testing}
-\section{Chi-Squared ($\chi^2$) Goodness of Fit Testing}
+        \subsection{Power of a test}
+            The power of a test is the probability of rejecting the null hypothesis, when the null hypothesis is false.The power of a test is the probability of rejecting the null hypothesis, when the null hypothesis is false.
+
+        \subsection{Test for diffrence between means of two normal distributions}
+            (Including approximation)
+        \subsection{Test for population proportion and for mean of a poisson distribution}
+        \subsection{Test for diffrence between population proportions and means of a poisson distribution}
+        \subsection{Test $H_0: \rho = 0$ for a bivariate normal population}

### 2017-06-04 src/mstats3.tex
Added ms03 hypothesis testing.

@@
         \subsection{Power of a test}
-            The power of a test is the probability of rejecting the null hypothesis, when the null hypothesis is false.The power of a test is the probability of rejecting the null hypothesis, when the null hypothesis is false.
+            The power of a test is the probability of rejecting the null hypothesis, when the null hypothesis is false.
+The power of a test is the probability of rejecting the null hypothesis, when the null hypothesis is false.
+This can be expressed as:
+            \\
+            $$1 - P(\text{Type II Error})$$
+            \\
+            To find the probability of a Type II error, consider the following hypothesis:
+            \begin{align*}
+            H_0&: \mu = 50   \\
+            H_1&: \mu \leq 50
+            \end{align*}
+            To find the probability of a type II error, we need to find the values of the test statistic were we reject $H_0$.\\
+            \begin{align*}
+                -1.34 &= \dfrac{\bar{X}-50}{\dfrac{21}{\sqrt{36}}} \\
+                        &= 45.31
+            \end{align*}
+            So we reject $H_0$ when $\bar{X} \leq 45.31$. If $\mu = 43$ what is $P(\text{Type II error})$. So we need to find $P(\text{Accepting } H_0 | H_0 \text{ is false})$, thus:
+            \begin{align*}
+            P(\bar{X} > 45.31 | \mu = 43) &= 1 - \phi\left(\dfrac{45.31 - 43}{\dfrac{21}{\sqrt{36}}}\right) \\
+                                          &= 0.255
+            \end{align*}
+            So the power of the test would be $1 - 0.255 = 0.745$.
+
+            %  of accepting $H_0$ given the $\mu = 26$. If the signifcance level of the test is $0.09$ then the critical value would be $-1.34$, and we will let the $\sigma=21$ and $n=36$.
+
+            % Now we need to find the value of the test statistic in which we reject the null hypothesis:
+            % Si
 
         \subsection{Test for diffrence between means of two normal distributions}
-            (Including approximation)
-        \subsection{Test for population proportion and for mean of a poisson distribution}
-        \subsection{Test for diffrence between population proportions and means of a poisson distribution}
+            (See SS05 Test for diffrence in means) (Including normal approximations)
+
+        \subsection{Test for population proportion}
+            (See SS04 Test for proportion)
+
+        \subsection{Test for mean of a poisson distribution}
+            (See SS04 Poisson hypothesis test)
+
+        \subsection{Test for diffrence between population proportions and means of two poisson distribution}
+            We use a normal approximation to test the difference between population proportions and means of a two poisson distribution, pooling their variances when $H_0$ is true.
+            \\\\
+            For example if $p_1 = \dfrac{205}{250}$ and $p_2 = \dfrac{137}{250}$ and $H_0: p_1 = p_2$, then $\hat{p} = \dfrac{205+137}{250+250} = \dfrac{342}{500} = \dfrac{171}{250}$
+            \\\\
+            (Couldnt find any questions on this)
+
         \subsection{Test $H_0: \rho = 0$ for a bivariate normal population}
+            (See SS03 Testing the correlation coefficent)

### 2017-06-04 src/stats3.tex
Added table of contents.
Removed new pages and created files for MS03 and MS04.
Added ms03 hypothesis testing.

--- 2017-04-06 stats3.tex
@@
-\chapter*{Statistics 3}
+\cchapter{Statistics 3}
 
@@
 
-    \newpage
     \subsection{Contingency Tables}
@@
 
-    \newpage
     \subsection{Yates Correction}
@@
         \end{example}
-
-\newpage
 \section{Distribution free methods}
@@
 
-    \newpage
     \subsection{The Sign Test}
@@
 
-    \newpage
     \subsection{Wilcoxon Signed-rank Test}
@@
 
-    \newpage
     \subsection{Mann-Whitney U-Test}
@@
 
-    \newpage
     \subsection{Kruskal-Wallis Test}
@@
         \end{example}
-
-\newpage
 \section{Correlation}
-    \newpage
     \subsection{Spearman's Rank Correlation Coefficent}
@@
 
-    \newpage
-    \subsection{Testing the correction coefficient}
+    \subsection{Testing the correlation coefficient}
         We may be interested in conducting a test to show their is no correlation between two random variables.
